### 2018-06-19 references.tex
references.tex - contains links to all important papers

\documentclass[12pt,a4paper,final]{article}
\usepackage[latin1]{inputenc}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{hyperref}
\title{References}
\begin{document}
\begin{enumerate}
    \item On AISync
    \begin{itemize}
        \item ZhaMot2017 
        \url{http://iopscience.iop.org/article/10.1088/1361-6544/aa8fe7/meta}

        \item ZhaNisMot2017 
        \url{http://iopscience.iop.org/article/10.1088/1361-6544/aa8fe7/meta}
    \end{itemize}
\end{enumerate}
\end{document}

### 2018-06-20 notes.tex
notes.tex added

\documentclass[12pt,a4paper,final]{article}
\usepackage[latin1]{inputenc}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{enumitem}
\date{}
\title{Asymmetry induced synchronization in FitzHugh Nagumo oscillators}
\begin{document}
    \maketitle
\section*{Asymmetry induced Synchronization (AISync)}
\begin{itemize}
    \item Stable synchronization may be observed in networks of non-identical 
    oscillators because of (and not despite) the difference between the 
    oscillators.This phenomenon is named AISync.
    \item AISync mediated by delay-coupling is observed in networks of 
    Stuart-Landau oscillators.
    \item  Possible in networks of non-identical oscillator with arbitrary 
    differences, provided that they admit {\it at least} one common orbit when 
    coupled.
    \item The idea is oscillator heterogeneity can stabilize an otherwise 
    unstable state of complete synchronization on representative networks.
    \item In other words, AISync refers to the situation wherein identically 
    coupled oscillators synchronize stably only when the oscillator themselves 
    are non-identical.
    \item How to demonstrate synchronization in coupled oscillators? \\
     {\it Generalized master stability analysis}\\
     technical points -
     \begin{enumerate}[label=(\alph*)]
     \item time delay in coupling
     \item finding the finest simultaneous block diagonalization.
    \end{enumerate}
\end{itemize}

\section*{FitzHugh Nagumo (FTH) oscillator}
\begin{itemize}
    \item Two identically coupled FHN oscillator with one or two different 
    time delays.
    \item For two time dealay $\rightarrow$ extreme events
    \item Why should we see AISync in a network of FHN oscillators?
    \begin{enumerate}[label=(\alph*)]
    \item Single delays -- synchronization is observed (common orbit exists)
    \item Two delays -- trajectories converge
    \end{enumerate}
\end{itemize}

\textbf{Questions:}
\begin{itemize}
    \item Can we observe AISync in coupled FHN oscillators? 
    \item What network structure to choose?
    \item How about on a ring network with fractal connectivities?
    \item What happens on a network with non-symmetric structure? 
    \item Two delays may generate extreme events. What does it translates to 
    if AISync is present?
\end{itemize}

\end{document}

### 2018-07-04 notes.tex
notes updated

@@
     are non-identical.
-    \item How to demonstrate synchronization in coupled oscillators? \\
-     {\it Generalized master stability analysis}\\
-     technical points -
+\end{itemize}
+
+\noindent \textbf{Conditions for AISyc}
+	\begin{enumerate}[label=(\Alph*)]
+		\item No asymptotically stable synchronization for homogeneous oscillators.
+
+		\item There is a heterogeneous configuration which results in stable asymptotically stable synchronization. 
+	\end{enumerate}
+
+\noindent \textbf{Some technical points}\\
+{\it may not be necessary for the preliminary studies}\\
+{\it Generalized master stability analysis}
+
 \begin{enumerate}[label=(\alph*)]
-     \item time delay in coupling
+	\item time delay in coupling (not necessary for AISync)
 	\item finding the finest simultaneous block diagonalization.
 \end{enumerate}
-\end{itemize}
 
@@
 
+\section*{Next steps}
+\begin{enumerate}
+	\item Choose a version of the FTH oscillator to work with including parameter values. 
+	\item The basic idea is that ``all kinds of " oscillators under consideration should admit a common orbit.
+Here, we will take just two kinds of oscillators -- this means that consider oscillators with two different sets of parameters.
+	\item Choose a simple symmetric network.
+	\item What kind of coupling are we going to use? Possibilities are -- diffusive coupling, adjacency matrix coupling, and time-delay coupling  
+	\item Challenge - Generalized master stability analysis using the finest simultaneous block diagonalization.
+	\item The previous step will give maximal transverse Lyapunov exponent. Synchronization happens if this exponent is negative.
+
+\end{enumerate}
+
 \end{document}

### 2018-07-11 notes.tex
notes updated

@@
-\documentclass[12pt,a4paper,final]{article}
-\usepackage[latin1]{inputenc}
-\usepackage{amsmath}
-\usepackage{amsfonts}
-\usepackage{amssymb}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% Short Sectioned Assignment
+% LaTeX Template
+% Version 1.0 (5/5/12)
+%
+% This template has been downloaded from:
+% http://www.LaTeXTemplates.com
+%
+% Original author:
+% Frits Wenneker (http://www.howtotex.com)
+%
+% License:
+% CC BY-NC-SA 3.0 (http://creativecommons.org/licenses/by-nc-sa/3.0/)
+%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%----------------------------------------------------------------------------------------
+%	PACKAGES AND OTHER DOCUMENT CONFIGURATIONS
+%----------------------------------------------------------------------------------------
+
+\documentclass[paper=a4, fontsize=12pt]{scrartcl} % A4 paper and 11pt font size
+
+\usepackage[T1]{fontenc} % Use 8-bit encoding that has 256 glyphs
+\usepackage{fourier} % Use the Adobe Utopia font for the document - comment 
+%this line to return to the LaTeX default
+\usepackage[english]{babel} % English language/hyphenation
+\usepackage{amsmath,amsfonts,amsthm} % Math packages
+\usepackage[a4paper, total={6.5in, 9in}]{geometry}
+\font\myfont=cmr12 at 14pt
+\usepackage{sectsty} % Allows customizing section commands
+\allsectionsfont{\centering \normalfont\scshape} % Make all sections centered, 
+%the default font and small caps
+\date{}
 \usepackage{graphicx}
@@
 \begin{enumerate}[label=(\alph*)]
-	\item time delay in coupling (not necessary for AISync)
+	\item time delay in coupling (could be the key ingredient for AISync)
 	\item finding the finest simultaneous block diagonalization.
@@
 
+\section{Introduction}
+FitzHugh Nagumo Oscillator:
+\begin{align}
+\frac{du}{dt} &= - w - u(u-1)(u-\lambda) \nonumber \\
+\frac{dw}{dt} &= \epsilon(u-\delta w)
+\end{align}
+
+Properties:
+\begin{itemize}
+    \item Limit cycle exists for $\epsilon\delta + \lambda\leq0$.
+    \item Stable limit cycle for $\epsilon > 0$ and $\delta < 
+    \frac{\lambda}{\epsilon}$
+    \item Hopf bifurcation occurs when $\epsilon\delta + \lambda = 0$
+\end{itemize}
+\section{Procedure}
+\begin{itemize}
+    \item We can introduce an asymmetry parameter $h$ such that $\delta 
+    \rightarrow \delta - \frac{h}{\epsilon}$ and $\lambda \rightarrow 
+    \lambda + h$.
+    \item Therefore, we consider two kinds of FHN oscillator with parameters 
+    $(\delta,\epsilon, \lambda)$ and $(\delta', \epsilon, \lambda')$ with 
+    asymmetry parameter $h$ such that   \begin{align*}
+    \delta' &= \delta - \frac{h}{\epsilon} \\
+    \lambda'  &=\lambda + h
+    \end{align*}
+    \item But we have to be careful about the choice of parameters so that 
+    conditions for stable limit cycle is met.
+    \item 
+\end{itemize}
 \end{document}

### 2018-08-21 Syncarea_control/Syncarea_control.tex
sync area manuscript

%
%   This file is part of the APS files in the REVTeX 4 distribution.
%   Version 4.0 of REVTeX, August 2001
%
%   Copyright (c) 2001 The American Physical Society.
%
%   See the REVTeX 4 README file for restrictions and more information.
%
% TeX'ing this file requires that you have AMS-LaTeX 2.0 installed
% as well as the rest of the prerequisites for REVTeX 4.0
%
% See the REVTeX 4 README file
% It also requires running BibTeX. The commands are as follows:
%                                                               /Users/neelimamgupte/Desktop/iitm.eps                                                                                                                                                                                 
%  1)  latex apssamp.tex
%  2)  bibtex apssamp
%  3)  latex apssamp.tex
%  4)  latex apssamp.tex
%
%\documentclass[twocolumn,showpacs,preprintnumbers,amsmath,amssymb]{revtex4-1}
\documentclass[preprint,showpacs,preprintnumbers,amsmath,amssymb]{revtex4-1}

% Some other (several out of many) possibilities
%\documentclass[preprint,aps]{revtex4-1}
%\documentclass[preprint,aps,draft]{revtex4}
%\documentclass[prb]{revtex4-1}% Physical Review B
\usepackage{graphicx}% Include figure files
\usepackage{dcolumn}% Align table columns on decimal point
\usepackage{bm}% bold math
\usepackage{amsmath,empheq}
\usepackage{hyperref}
\usepackage{natbib}
\newcommand{\overbar}[1]{\mkern 1.5mu\overline{\mkern-1.5mu#1\mkern-1.5mu}\mkern 1.5mu}

\begin{document}
\title{A mechanism to control synchronization in coupled area-preserving maps}
\affiliation{Department of Physics, Indian Institute of Technology Madras, Chennai, 600036, India.}
\author{Swetamber Das}
\email{swetdas@physics.iitm.ac.in}
\affiliation{Department of Physics, Indian Institute of Technology Madras, Chennai, 600036, India.}
%\author{Neelima Gupte}
%\email{gupte@physics.iitm.ac.in}
%\affiliation{Department of Physics, Indian Institute of Technology Madras, Chennai, 600036, India.}

\begin{abstract}
\pacs{05.45.+b}
\end{abstract}
\maketitle

\section{Introduction} 
Low-dimensional Hamiltonian systems commonly exhibit mixed phase space i.e. regular structures and chaotic regions may co-exist at a given degree of nonlinearity. This mixed nature has interesting consequences for the transport properties of such systems, for instance, the existence of anomalous kinetics, L\'{e}vy processes and L\'{e}vy flights \cite{Klafter1994,Zaburdaev2015}, power law contributions to recurrence and other statistics, and the existence of dynamical traps \cite{Zaslavsky2002a,Zaslavsky2002b}. An interesting region of a mixed phase space is the interface between regular regions and chaotic sea \cite{Mackay1984, Easton1993}. The dynamics in these neighbourhoods are complex and not well understood \cite{Meiss2015}. The complexity arises from the fact that a chaotic trajectory spends an arbitrary long but finite time at the boundaries of regular islands before exiting to the chaotic sea. The intermittent tendency of chaotic trajectories to stay close to the regular boundaries is called the phenomenon of stickiness. A major consequence of the stickiness is  the existence of power law in the Poincar{\'e} recurrence times indicating algebraic decay for long times times rather than exponential decay expected for normal transport. Therefore, due to stickiness, even a small regular island can influence the global transport properties of the system and decay of correlations. The phenomenon has been of great interest and continues to be studied on theoretical level \cite{Altmann2006, Altmann2010,Livorati2012,Bunimovich2012, Kruger2015}. In addition, stickiness has found application in several problems such as particle advection in fluids \cite{Babiano1994,Tel2005}, transport in plasma fusion devices \cite{Szezech2012,Martins2014}, celestial mechanics \cite{Efthymiopoulos1999, Harsoula2010, Harsoula2016}.

A recent feature of stickiness has emerged from a connected quarter. In our recent work~\cite{Mahata2016}, we have looked at the effects of the mixed phase space on the synchronization in a system of two standard maps coupled in unidirectional drive-response configuration. We have shown that synchronization of chaotic trajectories of the drive and response maps typically occurs in the neighborhood of regular inlands as a consequence of stickiness in the region.  This is the first instance, as far as we are aware, where stickiness have been found to influence synchronization in coupled chaotic systems.  For a chaotic orbit, synchronization typically happens via an intermittent behavior in the phase difference of the drive and response maps.   The sticky neighbourhoods of a regular islands contain Dynamical traps in chaotic orbits. In such trapping regions in the phase space, parts of a chaotic trajectory are almost regular in time and allow for synchronization to occur. Such traps are characterized using the properties of the finite-time Lyapunov exponent \cite{Szezech2005}. In the context of our work, we will refer to these traps as synchronization traps. We further noted that the behavior of the synchronization traps can be analyzed in a more quantitative way by analyzing the location and stability properties of the periodic orbits in the vicinity of the locations where synchronization takes place.

Synchronization in coupled chaotic systems, coupled map lattices and networks 
have been studied in great details over the years. The fact however remains 
that the ability to control synchronization based on the dynamics of the 
process has not been given much attention, until 
recently~\cite{Grabow2011,Wang2016}. In realistic systems, the speed of 
processing of visual or  synchronization may be significant. For example, in 
neurosciences,  the speed of the visual and olfactory processing are 
interesting problems \cite{Thorpe1996,Uchida2003}. In this paper, we intend 
to develop a couple of 
mechanisms to control synchronization process to wit,  to increase and to 
decrease synchronization time based on the location of synchronization traps 
in the sticky neightborhood of regular islands in the phase space of the 
area-preserving system of the standard map. We demonstrate the mechanism at $K 
= 6$ where only two small regular islands exist.  The rest of the paper is 
organized in the following way: we explain the coupling scheme in 
sec.~\ref{sec:Pecora_Carroll}, the mechanism and demonstration of delay and 
advanced synchronization times in sec.~\ref{sec:delay} and 
sec.~\ref{sec:advanced} respectively, and the paper ends with conclusions in 
sec.~\ref{sec:conclusions} including discussion on limitations and 
implications.

\section{The Pecora Caroll's unidirectional coupling scheme}
\label{sec:Pecora_Carroll} 
The standard map is considered to be the prototypical example of a two-dimensional area-preserving map, and is given by:

\begin{empheq}[right=\empheqrbrace \mod 1]{align}
P_{n+1} = P_n +  \frac{K}{2\pi}\sin(2\pi Q_n) \nonumber\\
Q_{n+1} = P_{n+1} + Q_n 
\end{empheq}

\begin{figure}[b]
\includegraphics[scale=.6]{Standard_Map}
\caption{\label{fig:Standard_map} \footnotesize The phase space of the standard map for  random initial conditions from the uniform distribution on $[0,1]$,  at the parameter value K = 6.0.}
\end{figure}
Here the subscript $n$ denotes the discrete time and $K$ is the nonlinearity parameter. These equations typically describe the evolution of two canonical variables $P$ and $Q$ which correspond to the momentum and co-ordinate in the Poincar\'{e} section of a freely moving rotator with interleaved periodic kicks. This system represents the behavior of a variety of systems such as charged particle confinement in  mirror magnetic traps, particle dynamics in accelerator, comet dynamics in solar systems etc.  Two-dimensional  phase space plots  of the standard map for the parameter value $K = 6$ using $25$ initial conditions  are shown in Figs.~\ref{fig:Standard_map}.

We now synchronize two standard maps, using the Pecora-Carroll scheme of synchronization using drive-response coupling \cite{Pecora1990,Pecora2015}. 
This system was first devised to synchronize the chaotic trajectories dissipative chaotic dynamical systems. Under this unidirectional coupling scheme, we duplicate the given map and couple the original and the duplicated map in a drive-response configuration. This means that the drive map evolves freely but the evolution of the response map is dependent on the drive. In this case, the $P$ value of the response system is set to the $P$ value of the drive system, at each iterate.  Therefore, the coupled system is described by the following equations

\begin{minipage}[t]{0.5\textwidth}
\begin{empheq}[right=\empheqrbrace \mod 1]{align}\nonumber
\label{equ:drive}
P^d_{n+1} = P^d_n + \frac{K}{2\pi}\sin(2\pi Q^d_n)\nonumber\\
Q^d_{n+1} = P^d_{n+1} + Q^d_n \nonumber
\end{empheq}
\centering \textbf{(Drive)}

\end{minipage}
\begin{minipage}[t]{0.5\textwidth}
\begin{empheq}[right=\empheqrbrace \mod 1]{align}
P^r_{n+1} = P^d_n + \frac{K}{2\pi}\sin(2\pi Q^r_n) \nonumber \\
Q^r_{n+1} = P^r_{n+1} + Q^r_n \nonumber
\end{empheq}
\centering \textbf{(Response)}
\end{minipage}

\vspace{.5cm}

The initial values of $Q$ in the drive and response maps are chosen arbitrarily, whereas the $P$ values are identical. The system is said to reach complete synchronization when both the $Q$ values of the drive and response systems evolve asymptotically to identical values i.e. 
\begin{equation}
\lim_{t\rightarrow\infty}(Q^d_n - Q^r_n) = 0
\end{equation}

Synchronization time is the least value of the iterate say, $\tau$, where $\Delta Q = Q^d-Q^r$ vanishes and remains so for rest of the iterations. 
\begin{equation}
(Q^d_n - Q^r_n) = 0; \hfill  n \geq  \bf {\tau }
\end{equation}

In all of the computations reported in the work, two chaotic trajectories are considered to be synchronized to numerical accuracy, if the Euclidean distance between them is less than $10^{-5}$. An example is shown in Fig.~\ref{fig:sync_ex}. Here, we plot $\Delta Q = Q^d_n-Q^r_n$ for initial conditions $(P^d_0,Q^d_0,Q^r_0) = (0.569, 0.906,0.106)$ at $K = 6$. The synchronization time is found to be 570246.

\begin{figure}[t]
\includegraphics[scale=.6]{Sync_exmaple}
\caption{\label{fig:sync_ex} \footnotesize The variation of $\Delta Q = Q^d_n-Q^r_n$ of the coupled system at $K = 6$ with iterations $n$. Synchronization is considered to be achieved when $\Delta Q < 10^{-5}$ and remain so for the rest of the iterations. The iteration step $n$ at which synchronization occurs is called synchronization time $\tau$. For this case, $\tau = 570 246.$ }
\end{figure}

We now analyze synchronization of the coupled system using the master stability function~\cite{Pecora1998}. The numerical procedures to control synchronization times will be discussed thereafter.

\section{Locating Synchronzation Traps using Master stability function}
\label{Master stability function}
A general drive-response system coupled unidirectionally may be described by the following set of equations

\begin{align}
\frac{d\overbar{X}_d}{dt} = \overbar{F}(\overbar{X}_d) \nonumber  &&
\frac{d\overbar{X_r}}{dt} = \overbar{F}(\overbar{X}_r) + \alpha E(\overbar{X}_d-\overbar{X}_r)
\end{align}

Here $\overbar{X}_d$ and $\overbar{X}_r$ are drive and response variables; the matrix $E$ determines the linear combination of the $\overbar{X}$ used in the difference and $\alpha$ is the coupling strength. For the map case, we have the following form

\begin{align}
\overbar{X}^d_{n+1} = \overbar{F}(\overbar{X}^d_{n}) \nonumber  &&
\overbar{X}^r_{n+1} = \overbar{F}(\overbar{X}^r_{n}) + \alpha E(\overbar{X}^d_{n}-\overbar{X}^r_n)
\end{align}

Therefore, in the case of a general unidirectional coupling of two standard maps, we get

\begin{center}
\begin{empheq}[right=\empheqrbrace \mod 1]{align}
P^d_{n+1} = P^d_n + \frac{K}{2\pi}\sin(2\pi Q^d_n) \nonumber\\
Q^d_{n+1} = P^d_{n+1} + Q^d_n \nonumber\\
P^r_{n+1} = P^r_n + \frac{K}{2\pi}\sin (2\pi Q^r_n) + \alpha(P^d_n-P^r_n) \nonumber \\
Q^r_{n+1} = P^r_{n+1} + Q^r_n 
\label{equ:coupled}
\end{empheq}
\end{center}

We have chosen $E$ to be the matrix $\begin{bmatrix} 1 & 0 \\ 1 & 0 \end{bmatrix}$.

To find the stability of the synchronous state, we first express  Eq.(\ref{equ:coupled}) in terms of  $P^\perp = P^d - P^r$ and $Q^\perp = Q^d - Q^r$, as follows
\begin{eqnarray}
\label{equ:trans}
P^\perp_{n+1} = (1-\alpha)P^\perp_n+ \frac{K}{2\pi} \sin(2\pi Q^d_n) - \frac{K}{2\pi} \sin(2\pi Q^r_n) \\ \nonumber
Q^\perp_{n+1} = (1-\alpha)P^\perp_n+ Q^\perp_n + \frac{K}{2\pi} \sin(2\pi Q^d_n) - \frac{K}{2\pi} \sin(2\pi Q^r_n)  
\end{eqnarray}

We now write the variational equation for Eq.(\ref{equ:trans}) by linearizing about $(P^d_n,Q^d_n)$
\begin{equation}
\label{equ:variational}
\left[ \begin{array}{c} \delta P^\perp_{n+1} \\ \delta Q^\perp_{n+1} \end{array} \right] = \mathcal{M}(\alpha)\left[ \begin{array}{c}\delta P^\perp_{n} \\ \delta Q^\perp_{n} \end{array} \right]
\end{equation}
where the matrix $\mathcal{M}(\alpha)$ is given by
\begin{equation} 
\label{equ:M}
\mathcal{M}(\alpha) = \begin{bmatrix} 1-\alpha & K\cos(2\pi Q^d_n) \\ 1 - \alpha & 1 + K\cos(2\pi Q^d_n) \end{bmatrix} 
\end{equation}

This is the master stability equation for the unidirectionally coupled standard map. The variational equation (\ref{equ:variational}) is the master stability equation for the coupled system under investigation. The associated largest LE computed from the master stability equation is the master stability function of the system, given by:
\begin{equation}
\label{equ:msf}
\lambda  = \lim_{n \rightarrow \infty} \lim_{\delta\overbar{X}_0\rightarrow 0}\frac{1}{n}\sum^{n-1}_{i=0}\ln|JM^n(\overbar{X}_i)|
\end{equation}

Here $n$ is a positive integer and $JM^n(\overbar{X}_i)$ denotes the Jacobian matrix of the $n$-times iterated map.  A negative value of the MSF (the largest non-zero LE) will ensure that $(P^\perp,Q^\perp) $ tend to zero indicating that the difference between $P$ and $Q$ will die out and the system will synchronize. 
Now, for the Pecora-Carroll approach, we set $\alpha = 1$. This substitution simplifies the matrix in Eq.(\ref{equ:M}) which now reads

\begin{equation} 
\mathcal{M}(1)= \begin{bmatrix} 0 & K\cos(2\pi Q^d_n) \\ 0 & 1 + K\cos(2\pi Q^d_n) \end{bmatrix} 
\end{equation}

The MSF should then be computed from the eigenvalues of the matrix $\mathcal{M}(1)$. It is easy to see that, for $\alpha =1$, one of the eigenvalues of   $\mathcal{M}(1)$ is zero. Therefore, we need to consider only the non-zero eigenvalue which is $1+K\cos(2\pi Q^d_n)$. The corresponding LE is given by

\begin{equation}
\label{equ:LE}
\lambda   = \lim_{n \rightarrow \infty} \frac{1}{n}\sum^{n-1}_{i=0}\ln|1+K\cos(2\pi Q^d_n)|
\end{equation}

In general, we define the $kth$ time-$n$ Lyapunov exponent associated with an initial point $\overbar{X}_0 = (P_0,Q_0)$ for a map $M(P,Q)$ as
\begin{equation}
\label{equ:FLEf}
\lambda _k(\overbar{X}_0;n)  = \frac{1}{n}\sum^{n-1}_{i=0}\ln|JM^n(\overbar{X}_i)|
\end{equation}

\begin{figure}[h]
\includegraphics[scale=0.55]{FTLE_plots}
\caption{\label{fig:FTLE}\footnotesize Fluctuations of time-5 FTLEs for three different set of initial conditions -- $(P^d_0,Q^d_0,Q^r_0) = (0.466,0.098,0.141)$ in (a),  $(0.009,0.026,0.129)$ in (b), and $(0.553,0.554,0.396)$ in (c). The synchronization times in each cases are 287090, 234521, and 164782 respectively. The time scale on the x-axis indicates iterates steps $t$ at which time-5 FTLEs have been computed. The drop window in (a), (b), and (c) shown in ellipses in red occur around synchronization times. Synchronization of chaotic trajectories in the coupled system under study typically occurs in one of such traps.}
\end{figure}

Here $n$ is a positive integer and $JM^n(\overbar{X}_i)$ denotes the Jacobian matrix of the $n$-times iterated map. We extend this notion to the master stability function defined in the Sec.~\ref{Master stability function}  i.e.  the non-zero Lyapunov exponent defined in Eq.(\ref{equ:LE}) 
which takes the following finite time version 
\begin{equation}
\lambda_1^s(\overbar{X}_0;n) = \frac{1}{n}\sum^{n-1}_{i=0}\ln|1 + K \cos (2\pi Q_i)|
\end{equation}

The subscript ($k = 1$) has been dropped hereafter as we have only one LE to compute.

We plot the variation in the time-5 FTLE defined above at $K = 6$. At the point of synchronization, the FTLE values attain a set of negative values consistently in a small window, indicating the existence of a synchronization trap.  In Fig.~\ref{fig:FTLE} , the three plots show the fluctuations  of FTLE near the point of synchronization for three different sets of initial conditions - $(P^d_0,Q^d_0,Q^r_0) = (0.466,0.098,0.141)$ in (a),  $(0.009,0.026,0.129)$ in (b), and $(0.553,0.554,0.396)$ in (c) with synchronization times 287090, 234521, and 164782 respectively. The values on the x-axis indicate time steps at which averages are computed so that synchronization times and the temporal neighborhood are effectively captured in the plots wherein a temporary drop is clearly visible, indicated by ellipses in red. Synchronization of chaotic trajectories typically occurs in one of such traps.

\section{Delayed Synchronization Times}
\label{sec:delay}
In order to develop a numerical technique to delay synchronization, we  first have to identify the vicinity of regular islands in the phase space. Another mechanism  to suppress stickiness based on knowledge of hyperbolic and non hyperbolic regions in the phase has been reported recently~\cite{Kruger2015}. Our method, however, targets the trajectories themselves in the sticky region to delay the process. We employ the edge-detection algorithm due to Benkadda {\it et al.} \cite{Benkadda1997}. For the numerical procedure, we divide the phase space in $100 \times 100$ grid. The edges of both the islands are detected by applying the standard map to point initiated in the chaotic sea, for $10^7$ times.  The edges thus detected in shown in Fig.  We have chosen the thickness parameter to be $d = 0.02$ i.e. euclidean distance of $d$ from points on the edge indicate the extent of the vicinity of the island.  This vicinity, therefore, indicate the the domain wherein the synchronization traps exist.  To demonstrate this explicitly, we compare the phase angles, defined by $\theta = \tan^{-1}\frac{Q}{P}$, of the points on the numerically detected edges and the points of synchronization, as follows.

In Fig.~\ref{fig:Edge_Sync_angles}(a), we show the distribution of phase angles of the points on the edges of regular islands determined by the edge-detection algorithm.  The bimodality of the distribution is due to the existence of two sharp islands in otherwise chaotic bulk in the phase space.  We compare this with the phase angles corresponding to the points in the phase space at which synchronization of the chaotic trajectories occurs.

\begin{figure}[h]
\includegraphics[scale=0.6]{Edge_Sync_angles}
%\includegraphics[scale=.4]{Standard_Map2.eps}\\
\caption{\label{fig:Edge_Sync_angles} \footnotesize Numerical detection of 
synchronization traps (a) Distribution of phase 
angles of the points on the edges of regular islands determined by the edge 
detection algorithm. (b) Distribution of phase angles for points where 
synchronization occurs. }
\end{figure}

Fig.~\ref{fig:Edge_Sync_angles}(b) shows the distribution of these phase angles for about 50 000 randomly chosen initial conditions for the drive and response maps that lead to synchronization. The locations of the peaks in the distribution shown in (a) approximately matches with this in (b). It is, therefore, visibly clear that synchronization occurs in the same angular domain of the phase space where the regular islands exist. Our numerically detected edges, thus correctly locate the synchronization traps in the phase space.  We now discuss the control mechanism to delay synchronization time.

The proximity parameter $d_0 = 0.02$ defines the region in the neighbourhood of the numerically detected edges wherein synchronization traps exist. The basic idea to control synchronization time is to identify the step at which the chaotic trajectory visits the numerically determined sticky neighbourhood followed by a slight deflection so that the trajectory resters at a point outside.  Our procedure depends upon how many times we deflect the trajectory  which will be referred to as step control. This means that if an $n-$step control is  employed then, the trajectory will be kicked away from the domain $n$-times during its first $n$ visits i.e. once per visit. The deflection is achieved by adding a randomly generated fraction below $0.1$ to the point visiting the domain, and therefore the maximum deflection area is roughly $1\%$ of the phase space.  This procedure applied on a given set of initial conditions of the drive and response maps, may result in four possibilities of synchronization time -- (1) successful delay  \textbf{(S)}, (2) no delay \textbf{(N)}, (3) failed to achieve synchronization \textbf{(F)}, and (4) undesirable faster synchronization \textbf{(U)}. Clearly, the only first possibility is the aim of the mechanism; the last one constitute a hazard while the other two are trivial. 
\begin{figure}[h]
\includegraphics[scale=0.6]{S_U_Percent}
\caption{\label{fig:Control_success} \footnotesize. The curves indicate the 
percentages of (S) successful delays shown as the blue curve,a and (U) 
undesirable fast synchronization showns as the red curve. }
\end{figure}

For the numerical implementation of the mechanism, we have considered $1000$ randomly chosen initial conditions from the chaotic sea of the phase space at $K = 6$. The maps have been iterated for $10^7$ times in each case.  The sticky domain is the determined by the proximity parameter $d_0$ around the numerically detected edges of the regular islands. We demonstrate the effectiveness of the control procedure  with respect to several $n-$step controls. The computations have been performed for $n \in \mathbb{A}$, where $\mathbb{A} = \{10, 20,30,...,90, 100, 150, ... ,10 000\}$.

\begin{figure}[t]
\includegraphics[scale=0.6]{Sync_time_dist.eps}
\caption{\label{fig:Sync_time_dist} \footnotesize. Distribution of 
synchronization time without n-step control (the blue curve) and with n-step 
control (the red curve). The control delays synchronization times. }
\end{figure}

Fig  shows the percentage of successful delays \textbf{(S)} (on the y-axis) for a given $n-$step control (on the x-axis) up to $n = 5500$ after which the successful cases are always $100\%$.  We achieve about $50\%$ successful cases for $n = 10$, which rises rapidly to $99\%$ for $n = 2950$. It is to be noted that the average synchronisation time without any control, $\tau_{avg} \sim 10^5$. Therefore, the n-steps required to obtain all successful delays is just about $3\%$ of $\tau_{avg}$. We also plot the distribution of synchronization times for successful cases and compare it against that of corresponding times without control in Fig. The long-tailed distribution of times without delay crosses over to a distribution which shows large flat region before a tail. The flat region indicate large number of successfully applied control with typical delay times of the order of $10^6$. For complexness, we show the decay in the number of undesired cases \textbf{U} in Fig.  The cases \textbf{F} and \textbf{N} which are largely present for $n<10$, are extremely small in number and are neglected.

 \begin{figure}[t]
\includegraphics[scale=0.6]{Strength_con.eps}
\caption{\label{fig:Strength_con} \footnotesize. Variation of $\beta$ 
(maximum, minimum and average) which measures the strength of successful 
delays i.e. $\beta = \ln(\frac{\tau_d}{\tau_0}$. $\beta_max$ indicates the 
maximum delay time obtained at a give n-step control while $\beta_{mean}$ 
stands for the mean value around 2. This means that on an average, 
synchronization times re delayed by 100 times. The minimum value of $\beta$ 
remains 1 for the cases when no delay was obtained obtained. }
\end{figure}

  The extent of delay $\beta$ may be defined as the ratio of successfully 
  delayed synchronization times $\tau_c$ to synchronization times $\tau_0$ 
  without control, i.e. $\beta =  \ln(\frac{\tau_d}{\tau_0}$).  Therefore, for 
  successfully delayed synchronization $\beta>0$ indicating $\tau_c > \tau$. 
  The plot in Fig.~\ref{fig:Strength_con} shows the variation of $\beta$ i.e. 
  $\beta_{min}$, $\beta_{mean}$, and $\beta_{max}$  indicating minimum, mean 
  and maximum $\beta$ for considered $n-$step controls. Clearly, we are able 
  to attain $\beta_{max}$ values to be in the range $5$ to $14$ while mean 
  values remain $\sim2$. Therefore, we are successful in achieving 
  significantly large delays in synchronization times.

There are a couple of limitations of the proposed control procedure. The deflection to the trajectory visiting the sticky domain should not insert it inside the regular island. The regular trajectories of the drive and response map are known to synchronize significantly faster and would result in the undesirable possibility \textbf{(U)}. Therefore, an estimate of the size of regular islands, in addition to their locations must be known. These may be determined efficiently, as we have shown, using the edge-detection algorithm. The deflection area  may then be chosen suitably.  Another limitation is that the procedure is unable to impose a pre-determined delay to synchronization times. The random deflections simply keep the trajectories temporarily away from the synchronization traps in the sticky domain and synchronization may eventually occur after the removal of the control.

 \section{Advanced Syncronization Times}
 \label{sec:advanced}
 We now explore the possibility to decrease synchronization times i.e. to expedite the process of synchronization. Once again, we will make use of the fact that synchronization traps typically exist in the sticky neighbourhoods of the regular islands.  The procedure is based on the parameter perturbation technique to generate coherent structures in the phase space of an area-preserving map. However, the technique may also be used to push the chaotic bulk on the coherent structure (regular region) already present in the phase space and therefore, to make it disappear completely.  We describe the procedure briefly.

 In a given area-preserving map, for instance, the  standard map, a  perturbation in the nonlinearity parameter $K$ in the neighbourhood of a suitable chosen periodic point leads to generation of large coherent structure.   We take the same form the standard map with a modulo operation such that $-0.5 \leq P_n \leq 0.5, 0\leq Q_n \leq 1$ . In the form above \cite{FN1}, the standard map is known to have a hyperbolic fixed point at $(0,0)$,$(0,0.5)$. We perturb the parameter $K$ to $K-\epsilon$ if $|P - P_f| < \delta$, $|Q-Q_f| < \delta$. For $P$ and $Q$ outside this $\delta$-strip, $K$ does not change.  It may be noted that the perturbed standard map remains to be area-preserving. The Jacobian $J$ is now given by:

\begin{minipage}[t]{0.55\textwidth}
%\raggedleft
\centering
\[ J = \left( \begin{array}{cc}
1 & 1 + (K-\epsilon)\cos(2\pi Q_n)   \\
1 & (K-\epsilon)\cos(2\pi Q_n) \end{array} \right)\]  
\end{minipage}
\begin{minipage}[t]{0.4\textwidth}
\raggedright

\vspace{0.65cm}
$ \epsilon \neq 0$ if $|P_f - P_n| < \delta$, $|Q_f - Q_n| < \delta$\\ $\epsilon = 0$, otherwise.\\
\end{minipage}

\vspace{0.65cm}
The determinant of this matrix $J$ remains 1. The phase space thus obtained for for $K = 6$ , $\delta = 0.3$ and $\epsilon = 2$ for the fixed point $(P_f,Q_f)=(0.0,0.0)$ is shown in Fig.~\ref{fig:coherent}.   We have performed computation for the other fixed point and with other modulo operations as well and results do not differ significantly. Our choice here of modulo operation and the fixed point $(P_f,Q_f)$ has some advantage over the others in terms of actual numerical values of synchronization times and the display of locations of the points where synchronization occurs.

For the computations, we take 50 000 initial conditions chosen randomly from the chaotic region of the unperturbed standard map.  The procedure has been applied to both drive and response map in accordance with our coupling  scheme described above. We choose a patch around $(P_f,Q_f)$  of length $\delta= 0.3$ which corresponds to about $9\%$ of the phase space area. We successfully obtain advance synchronization in more that $99.5\%$ of the cases among which a  couple of instances  has been shown in Fig.~\ref{fig:location} wherein synchronization occurs in the neighbourhood of the main regular island and one of its satellite islands.  The sign `+' in red indicate the point of synchronization. 
 \begin{figure}[h]
\includegraphics[scale=0.6]{Fast_sync_location.eps}
\caption{\label{fig:location} \footnotesize.The phase space of the drive map 
at K = 6 where parameter perturbation is applied. Synchronization occors at 
the location indicate by the red 'plus' signs.}
\end{figure}

 \begin{figure}[h]
\begin{tabular}{cc}
\includegraphics[scale=0.45]{Normal_Sync.eps}&
\includegraphics[scale=0.45]{Fast_sync_dist.eps}\\
(a) & (b)
\end{tabular}{}
\caption{\label{fig:sync_time_dist} \footnotesize The phase space of the standard map for 25 random initial conditions from the uniform distribution on $[0,2\pi]$ plotted for 4500 iterations; 500 iterates are discarded as transients. The values are normalized i.e. $\frac{P}{2\pi} $ and $\frac{Q}{2\pi}$ are plotted on the $y$ and $x$ axis for the parameter values (a)  K = 1.5, and (b) K = 6.0.}
\end{figure}

The probability distribution of synchronization times without parameter perturbation shows a long tail which shows a power law scaling. It is shown using cumulative probability distribution on a log-log scale in Fig.~\ref{fig:sync_time_dist}(a) with slope $\sim 0.744$. Upon application of perturbation, synchronization times become considerable shorter and the probability distribution cross over to exponential. Fig.~\ref{fig:sync_time_dist}(b) shows the resulting exponential distribution for the same set of initial conditions as shown for in Fig.~\ref{fig:sync_time_dist}(a) , with an estimated mean $\sim500$. Comparing the $x-$scales of the plots, one can see the faster synchronization has been successfully achieved.

We again define the strength of controlled synchronization as $\beta = \ln \frac{\tau_c}{\tau}$. For advanced synchronization, $\beta < 0$ indicating that synchronization times upon parameter perturbation is smaller than in normal course i.e. $\tau_c < \tau$. The probability distribution of $\beta$ in shown in Fig.~\ref{fig:beta_dist}, which may be fit with a gaussian of the following form:
\begin{equation}
f(\beta) =  a\exp(-\frac{(\beta-\mu)^2}{2c^2})
\end{equation}

We estimate the mean value $\mu \sim - 6.648$ and standard deviation $\sigma \sim1.542$ which corresponds to the fact that, on an average, advanced synchronization times thus obtained are $4\times10^2$ times smaller than the normal synchronization times with $95.5\%$ values are between $10$ to $10^4$ times smaller.  Therefore, the synchronization times we obtained upon parameter perturbation is significantly smaller than the normal synchronization times.

The parameter perturbation technique is an efficient way to decrease synchronization times with high accuracy.  The value of the perturbation imposed should be chosen suitably as a very small may not have any influence on the synchronization process. The procedure also depends on the size and location of the patch wherein the control is applied. We are yet to see if the other technique of generating coherent structures also influence synchronization in these unidirectionally coupled systems.

 \begin{figure}[h]
\includegraphics[scale=0.6]{Fast_strength_dist.eps}
\caption{\label{fig:location} \footnotesize.}
\end{figure}

\section{Conclusions}
\label{sec:conclusions}
To summarise, we have developed a numerical procedure to delay synchronization in two identical standard map coupled under the drive-response configuration. The procedure is based on the fact that synchronization in the system typically occurs in the sticky neighbourhoods of the regular islands in the phase space. We have shown that the delay is obtained by kicking the trajectories slightly away from the domain temporarily. The efficiency of the procedure depends upon the number of steps employed in control. By applying the procedure to the system on merely $1\%$ of the phase space area (at $K = 6$), we have successfully increased synchronization times. The maximum number of control steps that are used is $0.01\%$ of the total of number of iterations used throughout the computation.  Furthermore, the procedure gives variety of delays. The limitations of the procedure are, one, the deflection may insert the drive trajectory in the regular islands which would lead to undesirable faster synchronization and two, a predetermined synchronization time in not obtainable.

%\begin{thebibliography}
\bibliographystyle{unsrt}
\bibliography{Syncref}

%\end{thebibliography}
\end{document}

### 2018-08-23 Syncarea_control/coherent.py
coherent code

@@ -1,0 +1,133 @@
+from __future__ import division, print_function
+import numpy as np
+import matplotlib.animation as animation
+from scipy.integrate import odeint
+from numpy import arange
+from scipy.linalg import block_diag
+from CPG.hdf import HDFClassIO
+import CPG
+from pylab import *
+from sympy.physics.quantum.tensorproduct import TensorProduct
+def coherent_std_map(eps = 0.0, delta = 0.1, K = 6.0):
+
+    N = 10**7
+    n_incod = 50
+    del_Q = np.zeros(N)
+    P1_incod = np.random.uniform(-0.5,0.5,n_incod)
+    Q1_incod = np.random.uniform(0,1, n_incod)
+    Q2_incod = np.random.uniform(0,1,n_incod)
+
+    P_d = []
+    Q_d = []
+    Q_r = []
+    sync_time_control = []
+    for j in range(n_incod):
+        P1 = np.zeros(N)
+        P2 = np.zeros(N)
+        Q1 = np.zeros(N)
+        Q2 = np.zeros(N)
+        P1[0] = P1_incod[j]
+        Q1[0] = Q1_incod[j]
+        Q2[0] = Q2_incod[j]
+        for i in range(N-1):
+
+            if abs(P1[i]) < delta and abs(Q1[i]-0.5) < delta:
+                K1 = K - eps
+            else:
+                K1 = K
+            #print(K1)
+            P1[i+1] = P1[i] + K1/(2.0*np.pi)*np.sin(2.0*np.pi*(Q1[i]))
+            # P1[i+1] = (P1[i+1] + 0.5) % 1 - 0.5
+            Q1[i+1] = P1[i+1] + Q1[i]
+
+            if abs(P1[i]) < delta and abs(Q2[i]-0.5) < delta:
+                K2 = K - eps
+            else:
+                K2 = K
+
+            P2[i+1] = P1[i] + K2/(2.0*np.pi)*np.sin(2.0*np.pi*(Q2[i]))
+            Q2[i+1] = P2[i+1] + Q2[i]
+
+            P1[i+1] = (P1[i+1] + 0.5) % 1 - 0.5
+            Q1[i+1] = Q1[i+1] % 1
+            Q2[i+1] = Q2[i+1] % 1
+
+        diff = np.abs(Q1 - Q2)
+        # from IPython import embed
+        # embed()
+
+        for i in range(N-10):
+            if diff[i] < 10**-5:
+                # from IPython import embed
+                # embed()
[…]

### 2018-08-24 Syncarea_control/coherent_final.py
coherent code final.py

@@ -1,0 +1,121 @@
+from __future__ import division, print_function
+import numpy as np
+from numpy import arange
+from CPG.hdf import HDFClassIO
+import CPG
+import math
+
+def coherent_std_map(P1_incod, Q1_incod, Q2_incod, eps = 0.0, delta = 0.1, K = 6.0, N = 10, n_incod = 10):
+    sync_time = []
+    for j in range(n_incod):
+        P1 = np.zeros(N)
+        P2 = np.zeros(N)
+        Q1 = np.zeros(N)
+        Q2 = np.zeros(N)
+        P1[0] = P1_incod[j]
+        Q1[0] = Q1_incod[j]
+        Q2[0] = Q2_incod[j]
+        for i in range(N-1):
+            if abs(P1[i]) < delta and abs(Q1[i]) < delta:
+                K1 = K - eps
+            else:
+                K1 = K
+            P1[i+1] = P1[i] + K1/(2.0*math.pi)*math.sin(2.0*math.pi*Q1[i])
+            #P1[i+1] = (P1[i+1] + 0.5) % 1 - 0.5
+            Q1[i+1] = P1[i+1] + Q1[i]
+
+            if abs(P1[i]) < delta and abs(Q2[i]) < delta:
+                K2 = K - eps
+            else:
+                K2 = K
+            P2[i+1] = P1[i] + K2/(2.0*math.pi)*math.sin(2.0*math.pi*Q2[i])
+            Q2[i+1] = P2[i+1] + Q2[i]
+
+            if P1[i+1] > 0.5:
+                P1[i+1] = P1[i+1] - 1.0
+            elif P1[i+1] < -0.5:
+                P1[i+1] = P1[i+1] + 1.0
+            Q1[i+1] = Q1[i+1] % 1.0
+            Q2[i+1] = Q2[i+1] % 1.0
+            if i % 1000 == 0 and i != 0:
+               # print("i =",i)
+                # from IPython import embed
+                # embed()
+                diff = np.abs(Q1[i-1000:i] - Q2[i-1000:i])
+                if diff.max() < 10 ** -5:
+                    break
+        # P_d.append(P1)
+        # Q_d.append(Q1)
+        # Q_r.append(Q2)
+        # P_d.append(P1[N - 10000:N])
+        # Q_d.append(Q1[N - 10000:N])
+        # Q_r.append(Q2[N - 10000:N])
+        diff = np.abs(Q1 - Q2)
+        # from IPython import embed
[…]

### 2018-08-24 Syncarea_control/coherent_final.py
sync time code update

@@ -15,7 +15,8 @@ def coherent_std_map(P1_incod, Q1_incod, Q2_incod, eps = 0.0, delta = 0.1, K = 6
         P1[0] = P1_incod[j]
         Q1[0] = Q1_incod[j]
         Q2[0] = Q2_incod[j]
-        for i in range(N-1):
+        print("j =",j)
+	for i in range(N-1):
             if abs(P1[i]) < delta and abs(Q1[i]) < delta:
                 K1 = K - eps
             else:
@@ -63,12 +64,12 @@ def coherent_std_map(P1_incod, Q1_incod, Q2_incod, eps = 0.0, delta = 0.1, K = 6
                 if test.max() < 10**-5:
                     t = i
                     sync_time.append(t)
-                    print("sync time = ", i)
+                    #print("sync time = ", i)
 
                     break
         if t == 0:
             sync_time.append(t)
-            print("sync time = ", t)
+            #print("sync time = ", t)
 
     # plt.plot(Q_d,P_d,'.b')
     # plt.xlim(0,1)
@@ -84,7 +85,7 @@ def coherent_std_map(P1_incod, Q1_incod, Q2_incod, eps = 0.0, delta = 0.1, K = 6
 
 def sync_code(save_data = False):
     N = 10 ** 7
-    n_incod = 10
+    n_incod = 100
     del_Q = np.zeros(N)
     P1_incod = np.random.uniform(-0.4, 0.4, n_incod)
     Q1_incod = np.random.uniform(0.3, 0.7, n_incod)
@@ -97,11 +98,11 @@ def sync_code(save_data = False):
     print(sync_time_with_control)
     if save_data:
         data = dict(without_control=np.array(sync_time_without_control), with_control=np.array(sync_time_with_control))
-        fname = "sync_times_in_cod_50000.hf"
+        fname = "sync_times_in_cod_5000.hf"
         CPG.hdf.save_h5_dict(fname, data)
 
 def do_plots():
-    fname = "sync_times.hf"
+    fname = "sync_times_in_cod_5000.hf"
     data = CPG.hdf.load_h5_dict(fname)
 
     time_c = data["with_control"]
@@ -114,8 +115,8 @@ def do_plots():
 
 if __name__ == "__main__":
     #coherent_std_map(eps = 1.5, delta = 0.4, K = 6)
-    sync_code()
+    sync_code(save_data = "True")
     #test()
-    # do_plots()
+    #do_plots()
 
 

### 2018-08-24 Syncarea_control/coherent_final.py
code updated

@@ -5,8 +5,10 @@ from CPG.hdf import HDFClassIO
 import CPG
 import math
 
+
 def coherent_std_map(P1_incod, Q1_incod, Q2_incod, eps = 0.0, delta = 0.1, K = 6.0, N = 10, n_incod = 10):
     sync_time = []
+
     for j in range(n_incod):
         P1 = np.zeros(N)
         P2 = np.zeros(N)
@@ -15,8 +17,8 @@ def coherent_std_map(P1_incod, Q1_incod, Q2_incod, eps = 0.0, delta = 0.1, K = 6
         P1[0] = P1_incod[j]
         Q1[0] = Q1_incod[j]
         Q2[0] = Q2_incod[j]
-        print("j =",j)
-	for i in range(N-1):
+        #print("j =",j)
+        for i in range(N-1):
             if abs(P1[i]) < delta and abs(Q1[i]) < delta:
                 K1 = K - eps
             else:
@@ -85,7 +87,7 @@ def coherent_std_map(P1_incod, Q1_incod, Q2_incod, eps = 0.0, delta = 0.1, K = 6
 
 def sync_code(save_data = False):
     N = 10 ** 7
-    n_incod = 100
+    n_incod = 5000
     del_Q = np.zeros(N)
     P1_incod = np.random.uniform(-0.4, 0.4, n_incod)
     Q1_incod = np.random.uniform(0.3, 0.7, n_incod)
@@ -94,8 +96,8 @@ def sync_code(save_data = False):
     sync_time_without_control = coherent_std_map(N = N, n_incod = n_incod, eps = 0.0, delta = 0.0, K = 6, P1_incod = P1_incod, Q1_incod = Q1_incod, Q2_incod = Q2_incod)
     sync_time_with_control = coherent_std_map(N = N, n_incod = n_incod, eps = 1.5, delta = 0.4, K = 6, P1_incod = P1_incod, Q1_incod = Q1_incod, Q2_incod = Q2_incod)
 
-    print(sync_time_without_control)
-    print(sync_time_with_control)
+    #print(sync_time_without_control)
+    #print(sync_time_with_control)
     if save_data:
         data = dict(without_control=np.array(sync_time_without_control), with_control=np.array(sync_time_with_control))
         fname = "sync_times_in_cod_5000.hf"
@@ -115,7 +117,7 @@ def do_plots():
 
 if __name__ == "__main__":
[…]

### 2018-08-29 Syncarea_control/Syncarea_control.tex
sync area code + plot
syncarea tex updated

@@
 \author{Swetamber Das}
-\email{swetdas@physics.iitm.ac.in}
-\affiliation{Department of Physics, Indian Institute of Technology Madras, Chennai, 600036, India.}
+\email{swetdas@pks.mpg.de}
+\affiliation{Max Planck Institute for the Physics of Complex Systems \\ 
+N\"othnitzer Stra\ss e 38, Dresden, Germany.}
 %\author{Neelima Gupte}
@@
 \begin{abstract}
+    A unidirectionally coupled area-preserving maps with mixed phase space may 
+    show identical synchronization in the sticky neighborhood of the regular 
+    islands. We use this fact to devise a couple of numerical procedures to 
+    delay and expedite the process of synchronization in standard maps coupled 
+    under  Pecora-Caroll scheme. The delay method is based on controlled 
+    kicking of trajectories away from the regular islands for as long as 
+    necessary. On the other hand, the method to expiate the process is
+    achieved by a parameter perturbation technique which rapidly drives the 
+    chaotic trajectories to synchronization traps in the sticky neighborhoods 
+    of regular islands. We also point out of limitations of these methods. 
 \pacs{05.45.+b}

### 2018-08-29 Syncarea_control/paper.tex
paper updated

--- 2018-08-21 Syncarea_control/Syncarea_control.tex
@@
-%
-%   This file is part of the APS files in the REVTeX 4 distribution.
-%   Version 4.0 of REVTeX, August 2001
-%
-%   Copyright (c) 2001 The American Physical Society.
-%
-%   See the REVTeX 4 README file for restrictions and more information.
-%
-% TeX'ing this file requires that you have AMS-LaTeX 2.0 installed
-% as well as the rest of the prerequisites for REVTeX 4.0
-%
-% See the REVTeX 4 README file
-% It also requires running BibTeX. The commands are as follows:
-%                                                               /Users/neelimamgupte/Desktop/iitm.eps
-%  1)  latex apssamp.tex
-%  2)  bibtex apssamp
-%  3)  latex apssamp.tex
-%  4)  latex apssamp.tex
-%
-%\documentclass[twocolumn,showpacs,preprintnumbers,amsmath,amssymb]{revtex4-1}
-\documentclass[preprint,showpacs,preprintnumbers,amsmath,amssymb]{revtex4-1}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% **IMPORTANT** remarks ahead
+% - linelength <= 80 (in the text, unless unavoidable)
+% - no TABs
+% - no trailing spaces
+% - Use macros for more complicated variables
+%   (especially, if the notation might be changed later...)
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-% Some other (several out of many) possibilities
-%\documentclass[preprint,aps]{revtex4-1}
-%\documentclass[preprint,aps,draft]{revtex4}
-%\documentclass[prb]{revtex4-1}% Physical Review B
+\documentclass[reprint,superscriptaddress,amsmath,amssymb,aps,pre]{revtex4-1}
+
 \usepackage{graphicx}% Include figure files
+\graphicspath{plots}
 \usepackage{dcolumn}% Align table columns on decimal point
 \usepackage{bm}% bold math
-\usepackage{amsmath,empheq}
-\usepackage{hyperref}
-\usepackage{natbib}
+\usepackage{color}
+\usepackage{bbold}
+\usepackage{soul}
+
+\usepackage{empheq}
+\setcounter{topnumber}{1}
+
+\usepackage{flafter}     % Bilder immer nach \figure Befehl
+
+% http://tex.stackexchange.com/questions/62010/can-i-access-system-environment-variables-from-latex-for-instance-home
+\usepackage{catchfile}
+\newcommand{\getenv}[2][]{%
+\CatchFileEdef{\temp}{"|kpsewhich --var-value #2"}{\endlinechar=-1}%
+\if\relax\detokenize{#1}\relax\temp\else\let#1\temp\fi}
+
+%\getenv[\HOME]{HOME}
+
+% use okular to view this file and acroread will be called on links
+%\newcommand{\articlestoragepath}{/home/scratch/abaecker/ArticleStorage/}
+\getenv[\articlestoragepath]{ARTICLE_STORAGE}
+% Conditionally import pdf_article_link.sty to ensure that it works
+% even without this file.
+\IfFileExists{pdf_article_link.sty}{\usepackage{pdf_article_link}}
+
+\input defs.tex
 \newcommand{\overbar}[1]{\mkern 1.5mu\overline{\mkern-1.5mu#1\mkern-1.5mu}\mkern 1.5mu}
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% ADD THE FOLLOWING COUPLE LINES INTO YOUR PREAMBLE
+\let\OLDthebibliography\thebibliography
+\renewcommand\thebibliography[1]{
+  \OLDthebibliography{#1}
+  \setlength{\parskip}{0pt}
+  \setlength{\itemsep}{0pt plus 0.3ex}
+}
+
+\makeatletter
+\let\Hy@backout\@gobble
+\makeatother
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \begin{document}
-\title{A mechanism to control synchronization in coupled area-preserving maps}
-\affiliation{Department of Physics, Indian Institute of Technology Madras, Chennai, 600036, India.}
+
+\title{Mechanism to control synchronization in coupled area-preserving maps}
+
 \author{Swetamber Das}
-\email{swetdas@physics.iitm.ac.in}
-\affiliation{Department of Physics, Indian Institute of Technology Madras, Chennai, 600036, India.}
-%\author{Neelima Gupte}
-%\email{gupte@physics.iitm.ac.in}
-%\affiliation{Department of Physics, Indian Institute of Technology Madras, Chennai, 600036, India.}
+\affiliation{Max-Planck-Institut f\"ur Physik komplexer Systeme, N\"othnitzer
+    Stra\ss{}e 38, 01187 Dresden, Germany}
+
+\date{\today}
 
 \begin{abstract}
-\pacs{05.45.+b}
+A unidirectionally coupled area-preserving maps with mixed phase space may identical synchronization in the sticky neighborhood of the regular islands.
+We use this fact to devise numerical procedures to delay and advance the process of synchronization.
+The delay method is based on controlled kicking of the trajectories away from the regular islands for as long as required.
+The method to expedite the process is achieved by a parameter perturbation technique which rapidly drives the chaotic trajectories in the neighborhood of regular islands where synchronization occurs.
+We also point out the limitations of these methods.
 \end{abstract}
+
+\pacs{PACS here}
+
 \maketitle
@@
 Synchronization in coupled chaotic systems, coupled map lattices and networks 
-have been studied in great details over the years. The fact however remains 
+have been studied in great details over the years.
+Coupled Hamiltonian systems are also known to show synchronization such as measure synchronization~\cite{Hampton1999,Wang2003,Vincent2005,Gupta2017} and identical synchronization~\cite{Mahata2016,Das2017} The fact however remains
 that the ability to control synchronization based on the dynamics of the 
@@
 area-preserving system of the standard map. We demonstrate the mechanism at $K 
-= 6$ where only two small regular islands exist.  The rest of the paper is 
+= 6$ where only two small regular islands exist.
+It is important to note that role of stickiness in Hamiltonian systmes for synchronaization was already predicted by Zaslavsky~\cite{Zaslavsky2002b} early in the last decade.\\
+
+The rest of the paper is 
 organized in the following way: we explain the coupling scheme in 
@@
 \begin{figure}[b]
-\includegraphics[scale=.6]{Standard_Map}
+	\includegraphics[scale=.45]{Standard_Map}
 	\caption{\label{fig:Standard_map} \footnotesize The phase space of the standard map for  random initial conditions from the uniform distribution on $[0,1]$,  at the parameter value K = 6.0.}
@@
 \begin{figure}[t]
-\includegraphics[scale=.6]{Sync_exmaple}
+	\includegraphics[scale=.4]{Sync_exmaple}
 	\caption{\label{fig:sync_ex} \footnotesize The variation of $\Delta Q = Q^d_n-Q^r_n$ of the coupled system at $K = 6$ with iterations $n$.
@@
 
-\section{Locating Synchronzation Traps using Master stability function}
+\section{Locating Synchronization Traps using Master stability function}
 \label{Master stability function}
@@
 Therefore, in the case of a general unidirectional coupling of two standard maps, we get
-
-\begin{center}
-\begin{empheq}[right=\empheqrbrace \mod 1]{align}
-P^d_{n+1} = P^d_n + \frac{K}{2\pi}\sin(2\pi Q^d_n) \nonumber\\
-Q^d_{n+1} = P^d_{n+1} + Q^d_n \nonumber\\
-P^r_{n+1} = P^r_n + \frac{K}{2\pi}\sin (2\pi Q^r_n) + \alpha(P^d_n-P^r_n) \nonumber \\
-Q^r_{n+1} = P^r_{n+1} + Q^r_n 
+\begin{align}
+P^d_{n+1} &= P^d_n + \frac{K}{2\pi}\sin(2\pi Q^d_n) \nonumber\\
+Q^d_{n+1} &= P^d_{n+1} + Q^d_n \nonumber\\
+P^r_{n+1} &= P^r_n + \frac{K}{2\pi}\sin (2\pi Q^r_n) + \alpha(P^d_n-P^r_n) \nonumber \\
+Q^r_{n+1} &= P^r_{n+1} + Q^r_n 
 \label{equ:coupled}
-\end{empheq}
-\end{center}
+\end{align}
 
@@
 
-\begin{figure}[h]
-\includegraphics[scale=0.55]{FTLE_plots}
+\begin{figure*}[t!]
+	\includegraphics[scale=0.41]{FTLE_plots}
 	\caption{\label{fig:FTLE}\footnotesize Fluctuations of time-5 FTLEs for three different set of initial conditions -- $(P^d_0,Q^d_0,Q^r_0) = (0.466,0.098,0.141)$ in (a),  $(0.009,0.026,0.129)$ in (b), and $(0.553,0.554,0.396)$ in (c).
@@
 Synchronization of chaotic trajectories in the coupled system under study typically occurs in one of such traps.}
-\end{figure}
+\end{figure*}
 
@@
 \begin{figure}[h]
-\includegraphics[scale=0.6]{Edge_Sync_angles}
+	\includegraphics[scale=0.45]{Edge_Sync_angles}
 	%\includegraphics[scale=.4]{Standard_Map2.eps}\\
@@
 The proximity parameter $d_0 = 0.02$ defines the region in the neighbourhood of the numerically detected edges wherein synchronization traps exist.
-The basic idea to control synchronization time is to identify the step at which the chaotic trajectory visits the numerically determined sticky neighbourhood followed by a slight deflection so that the trajectory resters at a point outside.
+The basic idea to control synchronization time is to identify the step at which the chaotic trajectory visits the numerically determined sticky neighborhood followed by a slight deflection so that the trajectory restarts at a point outside.
 Our procedure depends upon how many times we deflect the trajectory  which will be referred to as step control.
@@
 \begin{figure}[h]
-\includegraphics[scale=0.6]{S_U_Percent}
-\caption{\label{fig:Control_success} \footnotesize. The curves indicate the 
+	\includegraphics[scale=0.45]{S_U_Percent}
+	\caption{\label{fig:Control_success}\footnotesize The curves indicate the 
 		percentages of (S) successful delays shown as the blue curve,a and (U) 
-undesirable fast synchronization showns as the red curve. }
+		undesirable fast synchronization shown as the red curve. }
 \end{figure}
@@
 \begin{figure}[t]
-\includegraphics[scale=0.6]{Sync_time_dist.eps}
-\caption{\label{fig:Sync_time_dist} \footnotesize. Distribution of 
+	\includegraphics[scale=0.45]{Sync_time_dist.eps}
+	\caption{\label{fig:Sync_time_dist}\footnotesize Distribution of 
 		synchronization time without n-step control (the blue curve) and with n-step 
@@
 \begin{figure}[t]
-\includegraphics[scale=0.6]{Strength_con.eps}
+	\includegraphics[scale=0.45]{Strength_con.eps}
 	\caption{\label{fig:Strength_con} \footnotesize. Variation of $\beta$ 
@@
 
-  The extent of delay $\beta$ may be defined as the ratio of successfully 
-  delayed synchronization times $\tau_c$ to synchronization times $\tau_0$ 
-  without control, i.e. $\beta =  \ln(\frac{\tau_d}{\tau_0}$).  Therefore, for 
-  successfully delayed synchronization $\beta>0$ indicating $\tau_c > \tau$. 
+The degree of delay $\beta$ may be defined as the ratio of synchronization times without numerical control, $\tau_c$ to synchronization times $\tau_0$ without control, i.e.
+$\beta =  \ln(\frac{\tau_d}{\tau_0}$).
+Therefore, for successfully delayed synchronization $\beta>0$ indicating $\tau_c > \tau$.
 The plot in Fig.~\ref{fig:Strength_con} shows the variation of $\beta$ i.e. 
@@
 
- \section{Advanced Syncronization Times}
+\section{Advanced Synchronization Times}
 \label{sec:advanced}
- We now explore the possibility to decrease synchronization times i.e.
+Based on the now explore the possibility to decrease synchronization times i.e.
 to expedite the process of synchronization.
@@
 
-\begin{minipage}[t]{0.55\textwidth}
+\begin{minipage}[t]{0.45\textwidth}
 	%\raggedleft
@@
 \end{minipage}
-\begin{minipage}[t]{0.4\textwidth}
-\raggedright
 
-\vspace{0.65cm}
-$ \epsilon \neq 0$ if $|P_f - P_n| < \delta$, $|Q_f - Q_n| < \delta$\\ $\epsilon = 0$, otherwise.\\
-\end{minipage}
+\vspace{0.4cm}
+	where, $ \epsilon \neq 0$ if $|P_f - P_n| < \delta$, $|Q_f - Q_n| < \delta$\\ $\epsilon = 0$, otherwise.\\
 
-\vspace{0.65cm}
+\begin{figure*}[t]
+	\includegraphics[scale=0.4]{Fast_sync_location.eps}
+	\caption{\label{fig:location}\footnotesize The phase space of the drive map at K = 6 for a couple of initial conditions where parameter perturbation is applied.
+Synchronization occurs at the location indicate by the red 'plus' signs.}
+\end{figure*}
+
 The determinant of this matrix $J$ remains 1.
@@
 
-For the computations, we take 50 000 initial conditions chosen randomly from the chaotic region of the unperturbed standard map.
+For the computations, we take 5 000 initial conditions chosen randomly from the chaotic region of the unperturbed standard map i.e.
+$P_0 \in \{-0.4,0.4\}$ and $Q_0 \in \{0.3,0.7\}$.
 The procedure has been applied to both drive and response map in accordance with our coupling  scheme described above.
-We choose a patch around $(P_f,Q_f)$  of length $\delta= 0.3$ which corresponds to about $9\%$ of the phase space area.
-We successfully obtain advance synchronization in more that $99.5\%$ of the cases among which a  couple of instances  has been shown in Fig.~\ref{fig:location} wherein synchronization occurs in the neighbourhood of the main regular island and one of its satellite islands.
+We choose a patch around $(P_f,Q_f)$  of length $\delta= 0.4$ which corresponds to about $16\%$ of the phase space area.
+We successfully obtain advance synchronization in more that $99.5\%$ of the cases among which a  couple of instances  has been shown in Fig.~\ref{fig:location} -- synchronization occurs in the neighborhood of the main regular island and one of its satellite islands.
 The sign `+' in red indicate the point of synchronization.
- \begin{figure}[h]
-\includegraphics[scale=0.6]{Fast_sync_location.eps}
-\caption{\label{fig:location} \footnotesize.The phase space of the drive map 
-at K = 6 where parameter perturbation is applied. Synchronization occors at 
-the location indicate by the red 'plus' signs.}
-\end{figure}
 
- \begin{figure}[h]
-\begin{tabular}{cc}
-\includegraphics[scale=0.45]{Normal_Sync.eps}&
-\includegraphics[scale=0.45]{Fast_sync_dist.eps}\\
-(a) & (b)
-\end{tabular}{}
-\caption{\label{fig:sync_time_dist} \footnotesize The phase space of the standard map for 25 random initial conditions from the uniform distribution on $[0,2\pi]$ plotted for 4500 iterations; 500 iterates are discarded as transients.
-The values are normalized i.e.
-$\frac{P}{2\pi} $ and $\frac{Q}{2\pi}$ are plotted on the $y$ and $x$ axis for the parameter values (a)  K = 1.5, and (b) K = 6.0.}
-\end{figure}
-
-The probability distribution of synchronization times without parameter perturbation shows a long tail which shows a power law scaling.
-It is shown using cumulative probability distribution on a log-log scale in Fig.~\ref{fig:sync_time_dist}(a) with slope $\sim 0.744$.
-Upon application of perturbation, synchronization times become considerable shorter and the probability distribution cross over to exponential.
-Fig.~\ref{fig:sync_time_dist}(b) shows the resulting exponential distribution for the same set of initial conditions as shown for in Fig.~\ref{fig:sync_time_dist}(a) , with an estimated mean $\sim500$.
-Comparing the $x-$scales of the plots, one can see the faster synchronization has been successfully achieved.
-
-We again define the strength of controlled synchronization as $\beta = \ln \frac{\tau_c}{\tau}$.
+We again define the degree of controlled synchronization as $\beta = \ln \frac{\tau_c}{\tau}$.
 For advanced synchronization, $\beta < 0$ indicating that synchronization times upon parameter perturbation is smaller than in normal course i.e.
@@
 \begin{equation}
-f(\beta) =  a\exp(-\frac{(\beta-\mu)^2}{2c^2})
+f(\beta) =  a\exp\Big(-\frac{(\beta-\mu)^2}{2c^2}\Big)
 \end{equation}
 
-We estimate the mean value $\mu \sim - 6.648$ and standard deviation $\sigma \sim1.542$ which corresponds to the fact that, on an average, advanced synchronization times thus obtained are $4\times10^2$ times smaller than the normal synchronization times with $95.5\%$ values are between $10$ to $10^4$ times smaller.
+We estimate the mean value $\mu \sim -7.46$ and standard deviation $\sigma \sim1.67$ which corresponds to the fact that, on an average, advanced synchronization times thus obtained are about $10^3$ times smaller than the normal synchronization times with $95.5\%$ values are between $10$ to $10^5$ times smaller.
 Therefore, the synchronization times we obtained upon parameter perturbation is significantly smaller than the normal synchronization times.
@@
 We are yet to see if the other technique of generating coherent structures also influence synchronization in these unidirectionally coupled systems.
-
- \begin{figure}[h]
-\includegraphics[scale=0.6]{Fast_strength_dist.eps}
-\caption{\label{fig:location} \footnotesize.}
+\begin{figure}[b]
+	\includegraphics[scale=0.45]{sync_time_dist_beta.pdf}
+	\caption{\label{fig:beta_dist}\footnotesize Distribution of degree of synchronization, $P(\beta)$.}
 \end{figure}
@@
 \label{sec:conclusions}
-To summarise, we have developed a numerical procedure to delay synchronization in two identical standard map coupled under the drive-response configuration.
-The procedure is based on the fact that synchronization in the system typically occurs in the sticky neighbourhoods of the regular islands in the phase space.
-We have shown that the delay is obtained by kicking the trajectories slightly away from the domain temporarily.
+To summarize, we have developed numerical procedures to control synchronization in two identical standard map coupled under the drive-response configuration.
+The procedure is based on the fact that synchronization in the system typically occurs in the sticky neighborhoods of the regular islands in the phase space.
+We have shown that a delay in synchronization can be obtained by kicking the trajectories slightly away from the domain temporarily.
 The efficiency of the procedure depends upon the number of steps employed in control.
@@
 The limitations of the procedure are, one, the deflection may insert the drive trajectory in the regular islands which would lead to undesirable faster synchronization and two, a predetermined synchronization time in not obtainable.
+Furthermore, we have demonstrated a simple way to fasten synchronization by the parameter perturbation technique which rapidly drives the chaotic trajectories in the sticky neighborhood of regular islands.
+The distribution of degree of synchronization times indicates the efficiency of the procedure.
+A possible future direction is to investigate the role of stickiness in measure synchronization in coupled Hamiltonian systems.
 
-%\begin{thebibliography}
-\bibliographystyle{unsrt}
+\begin{acknowledgments}
+I would like to thank Neelima Gupte and Arnd B\"acker for their suggestions to improve the manuscript. 
+\end{acknowledgments}
+
+\appendix
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%\section{Some stuff for the appendix}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%\bibliographystyle{cpg_unsrt_title}
+%\bibliographystyle{cpg_unsrt_key_title}
+%\bibliographystyle{cpg_unsrt_title2}
+%\bibliographystyle{cpg_unsrt_key_title2}
+%\bibliographystyle{unsrt}
+\bibliographystyle{cpg_unsrt_title_for_phys_rev}
+%\bibliographystyle{cpg_unsrt_key_title2} % to have labels + links
+
+% See ~/Devel/Repositories/Bibtex/notes.txt
+% if you need more
 \bibliography{Syncref}
 
-%\end{thebibliography}
 \end{document}

### 2018-08-29 Syncarea_control/paper.tex
paper updated

@@
 \begin{abstract}
-A unidirectionally coupled area-preserving maps with mixed phase space may identical synchronization in the sticky neighborhood of the regular islands.
-We use this fact to devise numerical procedures to delay and advance the process of synchronization.
-The delay method is based on controlled kicking of the trajectories away from the regular islands for as long as required.
-The method to expedite the process is achieved by a parameter perturbation technique which rapidly drives the chaotic trajectories in the neighborhood of regular islands where synchronization occurs.
-We also point out the limitations of these methods.
+ A unidirectionally coupled area-preserving maps with mixed phase space may 
+show identical synchronization in the sticky neighborhood of the regular 
+islands. We use this fact to devise a couple of numerical procedures to 
+delay and expedite the process of synchronization in standard maps coupled 
+under  Pecora-Caroll scheme. The delay method is based on controlled 
+kicking of trajectories away from the regular islands for as long as 
+necessary. On the other hand, the method to expiate the process is
+achieved by a parameter perturbation technique which rapidly drives the 
+chaotic trajectories to synchronization traps in the sticky neighborhoods 
+of regular islands. We also point out of limitations of these methods.  
 \end{abstract}

### 2018-08-30 Syncarea_control/paper.tex
paper updated

@@
 delay and expedite the process of synchronization in standard maps coupled 
-under  Pecora-Caroll scheme. The delay method is based on controlled 
-kicking of trajectories away from the regular islands for as long as 
+under  Pecora-Caroll coupling scheme. The delay method is based on controlled 
+kicking of trajectories away from synchronization traps for as long as 
 necessary. On the other hand, the method to expiate the process is
@@
 chaotic trajectories to synchronization traps in the sticky neighborhoods 
-of regular islands. We also point out of limitations of these methods.  
+of regular islands. We also point out of the limitations of these methods.
 \end{abstract}
@@
 Low-dimensional Hamiltonian systems commonly exhibit mixed phase space i.e. 
-regular structures and chaotic regions may co-exist at a given degree of nonlinearity.
-This mixed nature has interesting consequences for the transport properties of such systems, for instance, the existence of anomalous kinetics, L\'{e}vy processes and L\'{e}vy flights \cite{Klafter1994,Zaburdaev2015}, power law contributions to recurrence and other statistics, and the existence of dynamical traps \cite{Zaslavsky2002a,Zaslavsky2002b}.
-An interesting region of a mixed phase space is the interface between regular regions and chaotic sea \cite{Mackay1984, Easton1993}.
-The dynamics in these neighbourhoods are complex and not well understood \cite{Meiss2015}.
-The complexity arises from the fact that a chaotic trajectory spends an arbitrary long but finite time at the boundaries of regular islands before exiting to the chaotic sea.
-The intermittent tendency of chaotic trajectories to stay close to the regular boundaries is called the phenomenon of stickiness.
-A major consequence of the stickiness is  the existence of power law in the Poincar{\'e} recurrence times indicating algebraic decay for long times times rather than exponential decay expected for normal transport.
-Therefore, due to stickiness, even a small regular island can influence the global transport properties of the system and decay of correlations.
-The phenomenon has been of great interest and continues to be studied on theoretical level \cite{Altmann2006, Altmann2010,Livorati2012,Bunimovich2012, Kruger2015}.
-In addition, stickiness has found application in several problems such as particle advection in fluids \cite{Babiano1994,Tel2005}, transport in plasma fusion devices \cite{Szezech2012,Martins2014}, celestial mechanics \cite{Efthymiopoulos1999, Harsoula2010, Harsoula2016}.
+regular structures and chaotic regions may co-exist at a given degree of 
+nonlinearity. This mixed nature has interesting consequences for the transport
+properties of such systems, for instance, the existence of anomalous kinetics,
+L\'{e}vy processes and L\'{e}vy flights \cite{Klafter1994,Zaburdaev2015}, 
+power law contributions to recurrence and other statistics, and the existence 
+of dynamical traps \cite{Zaslavsky2002a,Zaslavsky2002b}. An interesting region 
+of a mixed phase space is the interface between regular regions and chaotic 
+sea \cite{Mackay1984, Easton1993}. The dynamics in these neighbourhoods are 
+complex but fairly well understood \cite{Meiss2015} for lower dimensional 
+systems. The complexity arises from the fact that a chaotic trajectory spends 
+an arbitrary long but finite time at the boundaries of regular islands before 
+exiting to the chaotic sea. The intermittent tendency of chaotic trajectories 
+to stay close to the regular boundaries is called the phenomenon of 
+stickiness. A major consequence of the stickiness is  the existence of power 
+law in the Poincar{\'e} recurrence times indicating algebraic decay for long 
+times times rather than exponential decay expected for normal transport. 
+Therefore, due to stickiness, even a small regular island can influence the 
+global transport properties of the system and decay of correlations. The 
+phenomenon has been of great interest and continues to be studied on 
+theoretical level \cite{Altmann2006, Altmann2010,Livorati2012,Bunimovich2012, 
+Kruger2015}. In addition, stickiness has found application in several problems 
+such as particle advection in fluids \cite{Babiano1994,Tel2005}, transport in 
+plasma fusion devices \cite{Szezech2012,Martins2014}, celestial mechanics 
+\cite{Efthymiopoulos1999, Harsoula2010, Harsoula2016}.
 
-A recent feature of stickiness has emerged from a connected quarter.
-In our recent work~\cite{Mahata2016}, we have looked at the effects of the mixed phase space on the synchronization in a system of two standard maps coupled in unidirectional drive-response configuration.
-We have shown that synchronization of chaotic trajectories of the drive and response maps typically occurs in the neighborhood of regular inlands as a consequence of stickiness in the region.
-This is the first instance, as far as we are aware, where stickiness have been found to influence synchronization in coupled chaotic systems.
-For a chaotic orbit, synchronization typically happens via an intermittent behavior in the phase difference of the drive and response maps.
-The sticky neighbourhoods of a regular islands contain Dynamical traps in chaotic orbits.
-In such trapping regions in the phase space, parts of a chaotic trajectory are almost regular in time and allow for synchronization to occur.
-Such traps are characterized using the properties of the finite-time Lyapunov exponent \cite{Szezech2005}.
-In the context of our work, we will refer to these traps as synchronization traps.
-We further noted that the behavior of the synchronization traps can be analyzed in a more quantitative way by analyzing the location and stability properties of the periodic orbits in the vicinity of the locations where synchronization takes place.
+A recent feature of stickiness has emerged from a connected quarter. In our 
+recent work~\cite{Mahata2016}, we have looked at the effects of the mixed 
+phase space on the synchronization in a system of two standard maps coupled in 
+unidirectional drive-response configuration. We have shown that 
+synchronization of chaotic trajectories of the drive and response maps 
+typically occurs in the neighborhood of regular inlands as a consequence of 
+stickiness in the region.  This is the first instance, as far as we are aware, 
+where stickiness have been found to influence synchronization in coupled 
+chaotic systems.  For a chaotic orbit, synchronization typically happens via 
+an intermittent behavior in the phase difference of the drive and response 
+maps.   The sticky neighborhoods of a regular islands contain dynamical traps 
+in chaotic orbits. In such trapping regions in the phase space, parts of a 
+chaotic trajectory are almost regular in time and allow for synchronization to 
+occur. Such traps are characterized using the properties of the finite-time 
+Lyapunov exponent \cite{Szezech2005}. In the context of our work, we will 
+refer to these traps as synchronization traps. We further noted that the 
+behavior of the synchronization traps can be analyzed in a more quantitative 
+way by analyzing the location and stability properties of the periodic orbits 
+in the vicinity of the locations where synchronization takes place.
 
@@
 recently~\cite{Grabow2011,Wang2016}. In realistic systems, the speed of 
-processing of visual or  synchronization may be significant. For example, in 
+synchronization may be significant. For example, in
 neurosciences,  the speed of the visual and olfactory processing are 
@@
 mechanisms to control synchronization process to wit,  to increase and to 
-decrease synchronization time based on the location of synchronization traps 
-in the sticky neightborhood of regular islands in the phase space of the 
+decrease synchronization times based on the location of synchronization traps 
+in the sticky neighborhood of regular islands in the phase space of the 
 area-preserving system of the standard map. We demonstrate the mechanism at $K 
-= 6$ where only two small regular islands exist.
-It is important to note that role of stickiness in Hamiltonian systmes for synchronaization was already predicted by Zaslavsky~\cite{Zaslavsky2002b} early in the last decade.\\
+= 6$ where only two small regular islands exist. It is important to point out 
+that the role of stickiness in Hamiltonian systems for synchronization was 
+already predicted by Zaslavsky~\cite{Zaslavsky2002b} early in the last decade.
 
@@
 organized in the following way: we explain the coupling scheme in 
-sec.~\ref{sec:Pecora_Carroll}, the mechanism and demonstration of delay and 
-advanced synchronization times in sec.~\ref{sec:delay} and 
-sec.~\ref{sec:advanced} respectively, and the paper ends with conclusions in 
-sec.~\ref{sec:conclusions} including discussion on limitations and 
-implications.
+sec.~\ref{sec:Pecora_Carroll} and the mechanism based on the location of 
+synchronization traps is given in sec.~\ref{Master stability function}. The 
+numerical procedures to delay and advanced synchronization times are 
+demonstrated in sec.~\ref{sec:delay} and sec.~\ref{sec:advanced} respectively, 
+and the paper ends with conclusions in sec.~\ref{sec:conclusions} including 
+discussion on limitations and implications.
 
@@
 \label{sec:delay}
-In order to develop a numerical technique to delay synchronization, we  first have to identify the vicinity of regular islands in the phase space.
-Another mechanism  to suppress stickiness based on knowledge of hyperbolic and non hyperbolic regions in the phase has been reported recently~\cite{Kruger2015}.
-Our method, however, targets the trajectories themselves in the sticky region to delay the process.
-We employ the edge-detection algorithm due to Benkadda {\it et al.} \cite{Benkadda1997}.
-For the numerical procedure, we divide the phase space in $100 \times 100$ grid.
-The edges of both the islands are detected by applying the standard map to point initiated in the chaotic sea, for $10^7$ times.
-The edges thus detected in shown in Fig.
-We have chosen the thickness parameter to be $d = 0.02$ i.e.
-euclidean distance of $d$ from points on the edge indicate the extent of the vicinity of the island.
-This vicinity, therefore, indicate the the domain wherein the synchronization traps exist.
-To demonstrate this explicitly, we compare the phase angles, defined by $\theta = \tan^{-1}\frac{Q}{P}$, of the points on the numerically detected edges and the points of synchronization, as follows.
+In order to develop a numerical technique to delay synchronization, we  first 
+have to identify the vicinity of regular islands in the phase space. A 
+mechanism  to suppress stickiness based on the knowledge of hyperbolic and non 
+hyperbolic regions in the phase has been reported recently~\cite{Kruger2015}. 
+Our method, however, targets the trajectories themselves in the sticky region 
+to delay the process.  It is to be noted that the proposed procedure ignores 
+the details of a rather complex hierarchy of structures around regular 
+islands. We employ the edge-detection algorithm due to Benkadda {\it et al.} 
+\cite{Benkadda1997} to identify the edges of regular islands. For the 
+numerical procedure, we divide the phase space in $100 \times 100$ grid. The 
+edges of both the islands are detected by applying the standard map to point 
+initiated in the chaotic sea, for $10^7$ times.  The edges thus detected in 
+shown in Fig.  We have chosen the thickness parameter to be $d =0.02$ i.e. 
+euclidean distance of $d$ from points on the edge indicate the extent of the 
+vicinity of the island.  This vicinity, therefore, indicate the domain wherein 
+the synchronization traps exist.  To demonstrate this explicitly, we compare 
+the phase angles, defined by $\theta = \tan^{-1}(\frac{Q}{P}$), of the points 
+on 
+the numerically detected edges and the points of synchronization, as follows.
 
@@
 	\caption{\label{fig:Control_success}\footnotesize The curves indicate the 
-		percentages of (S) successful delays shown as the blue curve,a and (U) 
+		percentages of (S) successful delays shown as the blue curve, and (U) 
 		undesirable fast synchronization shown as the red curve. }
@@
 	\caption{\label{fig:Sync_time_dist}\footnotesize Distribution of 
-		synchronization time without n-step control (the blue curve) and with n-step 
+		synchronization times without n-step control (the blue curve) and with 
+		n-step 
 		control (the red curve). The control delays synchronization times. }
@@
 
-Fig  shows the percentage of successful delays \textbf{(S)} (on the y-axis) for a given $n-$step control (on the x-axis) up to $n = 5500$ after which the successful cases are always $100\%$.
-We achieve about $50\%$ successful cases for $n = 10$, which rises rapidly to $99\%$ for $n = 2950$.
-It is to be noted that the average synchronisation time without any control, $\tau_{avg} \sim 10^5$.
-Therefore, the n-steps required to obtain all successful delays is just about $3\%$ of $\tau_{avg}$.
-We also plot the distribution of synchronization times for successful cases and compare it against that of corresponding times without control in Fig.
-The long-tailed distribution of times without delay crosses over to a distribution which shows large flat region before a tail.
-The flat region indicate large number of successfully applied control with typical delay times of the order of $10^6$.
-For complexness, we show the decay in the number of undesired cases \textbf{U} in Fig.
-The cases \textbf{F} and \textbf{N} which are largely present for $n<10$, are extremely small in number and are neglected.
+Fig.~\ref{fig:Control_success} shows the percentage of successful delays 
+\textbf{(S)} (on the y-axis) for a given $n-$step control (on the x-axis) up 
+to $n = 5500$ after which the successful cases are always $100\%$.  We achieve 
+about $50\%$ successful cases for $n = 10$, which rises rapidly to $99\%$ for 
+$n = 2950$. It is to be noted that the average synchronization time without 
+any control, $\tau_{avg} \sim 10^5$. Therefore, the n-steps required to obtain 
+all successful delays is just about $3\%$ of $\tau_{avg}$. We also plot the 
+distribution of synchronization times for successful cases and compare it 
+against that of corresponding times without control in 
+Fig~\ref{fig:Sync_time_dist}. The long-tailed 
+distribution of times without delay crosses over to a distribution which shows 
+large flat region before a tail. The flat region indicate large number of 
+successfully applied control with typical delay times of the order of $10^6$. 
+For completeness, we show the decay in the number of undesired cases 
+\textbf{U} in Fig~\ref{fig:Control_success}.  The cases \textbf{F} and 
+\textbf{N} which are largely present for $n<10$, are extremely small in number 
+and are neglected.
 
@@
 	\includegraphics[scale=0.45]{Strength_con.eps}
-	\caption{\label{fig:Strength_con} \footnotesize. Variation of $\beta$ 
+	\caption{\label{fig:Strength_con}\footnotesize Variation of $\beta$ 
 		(maximum, minimum and average) which measures the strength of successful 
-		delays i.e. $\beta = \ln(\frac{\tau_d}{\tau_0}$. $\beta_max$ indicates the 
-		maximum delay time obtained at a give n-step control while $\beta_{mean}$ 
-		stands for the mean value around 2. This means that on an average, 
-		synchronization times re delayed by 100 times. The minimum value of $\beta$ 
-		remains 1 for the cases when no delay was obtained obtained. }
+		delays i.e. $\beta = \ln(\frac{\tau_d}{\tau_0}$). $\beta_{max}$ 
+		indicates the maximum delay time obtained at a give n-step control 
+		while $\beta_{mean}$ stands for the mean value around 2. This means 
+		that on an average, synchronization times re delayed by 100 times. The 
+		minimum value of $\beta$ remains 1 for the cases when no delay was 
+		obtained. }
 \end{figure}
@@
 \label{sec:advanced}
-Based on the now explore the possibility to decrease synchronization times i.e.
-to expedite the process of synchronization.
-Once again, we will make use of the fact that synchronization traps typically exist in the sticky neighbourhoods of the regular islands.
-The procedure is based on the parameter perturbation technique to generate coherent structures in the phase space of an area-preserving map.
-However, the technique may also be used to push the chaotic bulk on the coherent structure (regular region) already present in the phase space and therefore, to make it disappear completely.
-We describe the procedure briefly.
+Based on the now explore the possibility to decrease synchronization times 
+i.e. to expedite the process of synchronization. Once again, we will make use 
+of the fact that synchronization traps typically exist in the sticky 
+neighborhoods of the regular islands.  The procedure is based on the parameter 
+perturbation technique to generate coherent structures in the phase space of 
+an area-preserving map. However, the technique may also be used to push the 
+chaotic bulk on the coherent structure (regular region) already present in the 
+phase space and therefore, to make it disappear completely.  We describe the 
+procedure briefly.
 
-In a given area-preserving map, for instance, the  standard map, a  perturbation in the nonlinearity parameter $K$ in the neighbourhood of a suitable chosen periodic point leads to generation of large coherent structure.
-We take the same form the standard map with a modulo operation such that $-0.5 \leq P_n \leq 0.5, 0\leq Q_n \leq 1$ .
-In the form above \cite{FN1}, the standard map is known to have a hyperbolic fixed point at $(0,0)$,$(0,0.5)$.
-We perturb the parameter $K$ to $K-\epsilon$ if $|P - P_f| < \delta$, $|Q-Q_f| < \delta$.
-For $P$ and $Q$ outside this $\delta$-strip, $K$ does not change.
-It may be noted that the perturbed standard map remains to be area-preserving.
-The Jacobian $J$ is now given by:
+In a given area-preserving map, for instance, the  standard map, a  
+perturbation in the nonlinearity parameter $K$ in the neighborhood of a 
+suitable chosen periodic point leads to generation of large coherent 
+structure.   We take the same form the standard map with a modulo operation 
+such that $-0.5 \leq P_n \leq 0.5, 0\leq Q_n \leq 1$ . In this form, the 
+standard map is known to have a hyperbolic fixed point at 
+$(0,0)$,$(0,0.5)$. We perturb the parameter $K$ to $K-\epsilon$ if $|P - P_f| 
+< \delta$, $|Q-Q_f| < \delta$. For $P$ and $Q$ outside this $\delta$-strip, 
+$K$ does not change.  It may be noted that the perturbed standard map remains 
+to be area-preserving. The Jacobian $J$ is now given by:
 
@@
 	\includegraphics[scale=0.4]{Fast_sync_location.eps}
-	\caption{\label{fig:location}\footnotesize The phase space of the drive map at K = 6 for a couple of initial conditions where parameter perturbation is applied.
-Synchronization occurs at the location indicate by the red 'plus' signs.}
+	\caption{\label{fig:location}\footnotesize The phase space of the drive 
+	map at K = 6 for a couple of initial conditions where parameter 
+	perturbation is applied. Synchronization occurs at the location indicate 
+	by the red `plus' signs.}
 \end{figure*}
 
-The determinant of this matrix $J$ remains 1.
-The phase space thus obtained for for $K = 6$ , $\delta = 0.3$ and $\epsilon = 2$ for the fixed point $(P_f,Q_f)=(0.0,0.0)$ is shown in Fig.~\ref{fig:coherent}.
-We have performed computation for the other fixed point and with other modulo operations as well and results do not differ significantly.
-Our choice here of modulo operation and the fixed point $(P_f,Q_f)$ has some advantage over the others in terms of actual numerical values of synchronization times and the display of locations of the points where synchronization occurs.
+The determinant of this matrix $J$ remains 1. The phase space thus obtained 
+for for $K = 6$ , $\delta = 0.4$ and $\epsilon = 2$ for the fixed point 
+$(P_f,Q_f)=(0.0,0.0)$ is shown in Fig.~\ref{fig:location}.   We have also 
+performed computation for the other fixed point and with other modulo 
+operations as well and results do not differ significantly. Our choice here of 
+modulo operation and the fixed point $(P_f,Q_f)$ has some advantage over the 
+others in terms of actual numerical values of synchronization times and the 
+display of locations of the points where synchronization occurs.
 
-For the computations, we take 5 000 initial conditions chosen randomly from the chaotic region of the unperturbed standard map i.e.
-$P_0 \in \{-0.4,0.4\}$ and $Q_0 \in \{0.3,0.7\}$.
-The procedure has been applied to both drive and response map in accordance with our coupling  scheme described above.
-We choose a patch around $(P_f,Q_f)$  of length $\delta= 0.4$ which corresponds to about $16\%$ of the phase space area.
-We successfully obtain advance synchronization in more that $99.5\%$ of the cases among which a  couple of instances  has been shown in Fig.~\ref{fig:location} -- synchronization occurs in the neighborhood of the main regular island and one of its satellite islands.
-The sign `+' in red indicate the point of synchronization.
+For the computations, we take 5 000 initial conditions chosen randomly from 
+the chaotic region of the unperturbed standard map i.e. $P_0 \in \{-0.4,0.4\}$ 
+and $Q_0 \in \{0.3,0.7\}$. The total number of iterations used for each 
+initial condition to check for synchronization is $10^7$ and threshold of 
+synchronization is $10^{-5}$ as previously.  The procedure has 
+been applied to both drive and 
+response map in accordance with our coupling  scheme described above. We 
+choose a patch around $(P_f,Q_f)$  of length $\delta= 0.4$ which corresponds 
+to about $16\%$ of the phase space area. We successfully obtain advance 
+synchronization in more that $99.5\%$ of the cases among which a  couple of 
+instances  has been shown in Fig.~\ref{fig:location} -- synchronization occurs 
+in the neighborhood of one of the regular islands and one of their satellite 
+islands.  The sign `+' in red indicate the point of synchronization. For 
+details for the synchronization process, see ~\cite{Mahata2016,Das2017}.
 
-We again define the degree of controlled synchronization as $\beta = \ln \frac{\tau_c}{\tau}$.
-For advanced synchronization, $\beta < 0$ indicating that synchronization times upon parameter perturbation is smaller than in normal course i.e.
-$\tau_c < \tau$.
-The probability distribution of $\beta$ in shown in Fig.~\ref{fig:beta_dist}, which may be fit with a gaussian of the following form:
+We again define the degree of controlled synchronization as $\beta = \ln 
+\frac{\tau_c}{\tau}$. For advanced synchronization, $\beta < 0$ indicating 
+that synchronization times upon parameter perturbation is smaller than in 
+normal course i.e. $\tau_c < \tau$. The normalized probability distribution of 
+$\beta$ in shown in Fig.~\ref{fig:beta_dist}, which may be fit with a Gaussian 
+of the following form:
 \begin{equation}
-f(\beta) =  a\exp\Big(-\frac{(\beta-\mu)^2}{2c^2}\Big)
+f(\beta) =  a\exp\Big(-\frac{(\beta-\mu)^2}{2\sigma^2}\Big)
 \end{equation}
 
-We estimate the mean value $\mu \sim -7.46$ and standard deviation $\sigma \sim1.67$ which corresponds to the fact that, on an average, advanced synchronization times thus obtained are about $10^3$ times smaller than the normal synchronization times with $95.5\%$ values are between $10$ to $10^5$ times smaller.
-Therefore, the synchronization times we obtained upon parameter perturbation is significantly smaller than the normal synchronization times.
+We estimate the mean value $\mu \sim -7.46$ and standard deviation $\sigma 
+\sim1.67$ which corresponds to the fact that, on an average, advanced 
+synchronization times thus obtained are about $10^3$ times smaller than the 
+normal synchronization times with $95.5\%$ values are between $10$ to $10^5$ 
+times smaller.  Therefore, the synchronization times we obtained upon 
+parameter perturbation is significantly smaller than the normal 
+synchronization times.
 
@@
 \begin{figure}[b]
-	\includegraphics[scale=0.45]{sync_time_dist_beta.pdf}
-	\caption{\label{fig:beta_dist}\footnotesize Distribution of degree of synchronization, $P(\beta)$.}
+	%\includegraphics[scale=0.45]{sync_time_dist_beta.pdf}
+    \includegraphics[scale=0.45]{Fast_strength_dist.eps}
+	\caption{\label{fig:beta_dist}\footnotesize Distribution of degree of 
+	ynchronization, $P(\beta)$, mean is -7.46 and standard deviation 1.67.}
 \end{figure}
@@
 \label{sec:conclusions}
-To summarize, we have developed numerical procedures to control synchronization in two identical standard map coupled under the drive-response configuration.
-The procedure is based on the fact that synchronization in the system typically occurs in the sticky neighborhoods of the regular islands in the phase space.
-We have shown that a delay in synchronization can be obtained by kicking the trajectories slightly away from the domain temporarily.
-The efficiency of the procedure depends upon the number of steps employed in control.
-By applying the procedure to the system on merely $1\%$ of the phase space area (at $K = 6$), we have successfully increased synchronization times.
-The maximum number of control steps that are used is $0.01\%$ of the total of number of iterations used throughout the computation.
-Furthermore, the procedure gives variety of delays.
-The limitations of the procedure are, one, the deflection may insert the drive trajectory in the regular islands which would lead to undesirable faster synchronization and two, a predetermined synchronization time in not obtainable.
-Furthermore, we have demonstrated a simple way to fasten synchronization by the parameter perturbation technique which rapidly drives the chaotic trajectories in the sticky neighborhood of regular islands.
-The distribution of degree of synchronization times indicates the efficiency of the procedure.
-A possible future direction is to investigate the role of stickiness in measure synchronization in coupled Hamiltonian systems.
+To summarize, we have developed numerical procedures to control 
+synchronization in two identical standard map coupled under the drive-response 
+configuration. The procedure is based on the fact that synchronization in the 
+system typically occurs in the sticky neighborhoods of the regular islands in 
+the phase space. We have shown that a delay in synchronization can be obtained 
+by kicking the trajectories slightly away from the domain temporarily. The 
+efficiency of the procedure depends upon the number of steps employed in 
+control. By applying the procedure to the system on merely $1\%$ of the phase 
+space area (at $K = 6$), we have successfully increased synchronization times. 
+The maximum number of control steps that are used is $0.01\%$ of the total of 
+number of iterations used throughout the computation.  Furthermore, the 
+procedure gives variety of delays. The limitations of this procedure are, one, 
+the deflection may insert the drive trajectory in the regular islands which 
+would lead to undesirable faster synchronization and two, a predetermined 
+synchronization time in not obtainable. Furthermore, we have demonstrated a 
+simple way to fasten synchronization by the parameter perturbation technique 
+which rapidly drives the chaotic trajectories in the sticky neighborhood of 
+regular islands. The distribution of degree of synchronization times indicates 
+the efficiency of the procedure. In both the numerical procedures, we have 
+ignored the complex hierarchical structures which lead to sticky 
+behavior of the chaotic orbits in the phase space. A possible future direction 
+is to investigate the role of stickiness in measure synchronization in coupled 
+Hamiltonian systems.
 

### 2018-08-30 Syncarea_control/coherent_final.py
sync code updated

@@ -117,8 +117,8 @@ def do_plots():
 
 if __name__ == "__main__":
     #coherent_std_map(eps = 1.5, delta = 0.4, K = 6)
-    sync_code(save_data = True)
+    #sync_code(save_data = True)
     #test()
-    #do_plots()
+    do_plots()
 
 

### 2018-08-31 Syncarea_control/paper.tex
paper updated

@@
 islands. We employ the edge-detection algorithm due to Benkadda {\it et al.} 
-\cite{Benkadda1997} to identify the edges of regular islands. For the 
-numerical procedure, we divide the phase space in $100 \times 100$ grid. The 
-edges of both the islands are detected by applying the standard map to point 
-initiated in the chaotic sea, for $10^7$ times.  The edges thus detected in 
-shown in Fig.  We have chosen the thickness parameter to be $d =0.02$ i.e. 
-euclidean distance of $d$ from points on the edge indicate the extent of the 
-vicinity of the island.  This vicinity, therefore, indicate the domain wherein 
-the synchronization traps exist.  To demonstrate this explicitly, we compare 
-the phase angles, defined by $\theta = \tan^{-1}(\frac{Q}{P}$), of the points 
-on 
-the numerically detected edges and the points of synchronization, as follows.
+\cite{Benkadda1997} to identify the edges of regular islands in phase space. 
+For the numerical procedure, we divide the phase space in $100 \times 100$ 
+grid. The edges of both the islands are detected by applying the standard map 
+to point initiated in the chaotic sea, for $10^7$ times.  The edges thus 
+detected in shown in Fig~\ref{fig:edge_reg_island}.  We have chosen the 
+thickness parameter to be $d=0.02$ i.e. euclidean distance of $d$ from points 
+on the edge indicate the extent of the vicinity of the island.  This vicinity, 
+therefore, indicate the domain wherein the synchronization traps exist.  To 
+demonstrate this explicitly, we compare the phase angles, defined by $\theta = 
+\tan^{-1}(\frac{Q}{P}$), of the points on the numerically detected edges and 
+the points of synchronization, as 
+follows.
+
+\begin{figure}[b]
+    \includegraphics[scale=0.42]{edge_reg_island.eps}
+    %\includegraphics[scale=.4]{Standard_Map2.eps}\\
+    \caption{\label{fig:edge_reg_island} \footnotesize Example to detect the 
+    edge of a regular island in the phase space using edge detection 
+    algorithm  due to Benkadda {\it et al.} \cite{Benkadda1997}. }
+\end{figure}
 
@@
 
-The proximity parameter $d_0 = 0.02$ defines the region in the neighbourhood of the numerically detected edges wherein synchronization traps exist.
-The basic idea to control synchronization time is to identify the step at which the chaotic trajectory visits the numerically determined sticky neighborhood followed by a slight deflection so that the trajectory restarts at a point outside.
-Our procedure depends upon how many times we deflect the trajectory  which will be referred to as step control.
-This means that if an $n-$step control is  employed then, the trajectory will be kicked away from the domain $n$-times during its first $n$ visits i.e.
-once per visit.
-The deflection is achieved by adding a randomly generated fraction below $0.1$ to the point visiting the domain, and therefore the maximum deflection area is roughly $1\%$ of the phase space.
-This procedure applied on a given set of initial conditions of the drive and response maps, may result in four possibilities of synchronization time -- (1) successful delay  \textbf{(S)}, (2) no delay \textbf{(N)}, (3) failed to achieve synchronization \textbf{(F)}, and (4) undesirable faster synchronization \textbf{(U)}.
-Clearly, the only first possibility is the aim of the mechanism; the last one constitute a hazard while the other two are trivial.
+The proximity parameter $d_0 = 0.02$ defines the region in the neighborhood of 
+the numerically detected edges wherein synchronization traps exist. The basic 
+idea to control synchronization time is to identify the step at which the 
+chaotic trajectory visits the numerically determined sticky neighborhood 
+followed by a slight deflection so that the trajectory restarts at a point 
+outside.  Our procedure depends upon how many times we deflect the trajectory  
+which will be referred to as step control. This means that if an $n-$step 
+control is  employed then, the trajectory will be kicked away from the domain 
+$n$-times during its first $n$ visits i.e. once per visit. The deflection is 
+achieved by adding a randomly generated fraction below $0.1$ to the point 
+visiting the domain, and therefore the maximum deflection area is roughly 
+$1\%$ of the phase space.  This procedure applied on a given set of initial 
+conditions of the drive and response maps, may result in four possibilities of 
+synchronization time -- (1) successful delay  \textbf{(S)}, (2) no delay 
+\textbf{(N)}, (3) failed to achieve synchronization \textbf{(F)}, and (4) 
+undesirable faster synchronization \textbf{(U)}. Clearly, the only first 
+possibility is the aim of the mechanism; the last one constitute a hazard 
+while the other two are trivial. 
 \begin{figure}[h]
@@
 
+\begin{figure}[b]
+    \includegraphics[scale=0.45]{Strength_con.eps}
+    \caption{\label{fig:Strength_con}\footnotesize Variation of $\beta$ 
+        (maximum, minimum and average) which measures the strength of 
+        successful 
+        delays i.e. $\beta = \ln(\frac{\tau_d}{\tau_0}$). $\beta_{max}$ 
+        indicates the maximum delay time obtained at a give n-step control 
+        while $\beta_{mean}$ stands for the mean value around 2. This means 
+        that on an average, synchronization times re delayed by 100 times. The 
+        minimum value of $\beta$ remains 1 for the cases when no delay was 
+        obtained. }
+\end{figure}
+
 Fig.~\ref{fig:Control_success} shows the percentage of successful delays 
@@
 
-\begin{figure}[t]
-	\includegraphics[scale=0.45]{Strength_con.eps}
-	\caption{\label{fig:Strength_con}\footnotesize Variation of $\beta$ 
-		(maximum, minimum and average) which measures the strength of successful 
-		delays i.e. $\beta = \ln(\frac{\tau_d}{\tau_0}$). $\beta_{max}$ 
-		indicates the maximum delay time obtained at a give n-step control 
-		while $\beta_{mean}$ stands for the mean value around 2. This means 
-		that on an average, synchronization times re delayed by 100 times. The 
-		minimum value of $\beta$ remains 1 for the cases when no delay was 
-		obtained. }
-\end{figure}
-
-The degree of delay $\beta$ may be defined as the ratio of synchronization times without numerical control, $\tau_c$ to synchronization times $\tau_0$ without control, i.e.
-$\beta =  \ln(\frac{\tau_d}{\tau_0}$).
-Therefore, for successfully delayed synchronization $\beta>0$ indicating $\tau_c > \tau$.
-The plot in Fig.~\ref{fig:Strength_con} shows the variation of $\beta$ i.e. 
-$\beta_{min}$, $\beta_{mean}$, and $\beta_{max}$  indicating minimum, mean 
-and maximum $\beta$ for considered $n-$step controls. Clearly, we are able 
-to attain $\beta_{max}$ values to be in the range $5$ to $14$ while mean 
-values remain $\sim2$. Therefore, we are successful in achieving 
+The degree of controlled synchronization $\beta$ may be defined as the ratio 
+of synchronization times without numerical control, $\tau_c$ to 
+synchronization times $\tau_0$ without control, i.e. $\beta =  
+\ln(\frac{\tau_d}{\tau_0}$).  Therefore, for successfully delayed 
+synchronization $\beta>0$ indicating $\tau_c > \tau$.The plot in 
+Fig.~\ref{fig:Strength_con} shows the variation of $\beta$ i.e. $\beta_{min}$, 
+$\beta_{mean}$, and $\beta_{max}$  indicating minimum, mean and maximum 
+$\beta$ for considered $n-$step controls. Clearly, we are able to attain 
+$\beta_{max}$ values to be in the range $5$ to $14$ while mean values remain 
+$\sim2$. Therefore, we are successful in achieving 
 significantly large delays in synchronization times.
@@
 
+\begin{figure*}[t]
+    \includegraphics[scale=0.5]{Fast_sync_location.eps}
+    \caption{\label{fig:location}\footnotesize The phase space of the drive 
+        map at K = 6 for a couple of initial conditions where parameter 
+        perturbation is applied. Synchronization occurs at the location 
+        indicate 
+        by the red `plus' signs.}
+\end{figure*}
 In a given area-preserving map, for instance, the  standard map, a  
@@
 
-\vspace{0.4cm}
-	where, $ \epsilon \neq 0$ if $|P_f - P_n| < \delta$, $|Q_f - Q_n| < \delta$\\ $\epsilon = 0$, otherwise.\\
-
-\begin{figure*}[t]
-	\includegraphics[scale=0.4]{Fast_sync_location.eps}
-	\caption{\label{fig:location}\footnotesize The phase space of the drive 
-	map at K = 6 for a couple of initial conditions where parameter 
-	perturbation is applied. Synchronization occurs at the location indicate 
-	by the red `plus' signs.}
-\end{figure*}
+\vspace{0.2cm}
+	where, $ \epsilon \neq 0$ if $|P_f - P_n| < \delta$, $|Q_f - Q_n| < 
+	\delta$\\ $\epsilon = 0$, otherwise.\\
 
@@
 to about $16\%$ of the phase space area. We successfully obtain advance 
-synchronization in more that $99.5\%$ of the cases among which a  couple of 
+synchronization in more than $99.5\%$ of the cases among which a  couple of 
 instances  has been shown in Fig.~\ref{fig:location} -- synchronization occurs 
@@
 details for the synchronization process, see ~\cite{Mahata2016,Das2017}.
-
+\begin{figure}[h]
+    \includegraphics[scale=0.8]{sync_time_beta.eps}
+    \caption{\label{fig:beta_dist}\footnotesize Distribution of degree of 
+        controlled synchronization, $P(\beta)$, mean is -7.46 and standard 
+        deviation is 1.67.}
+\end{figure}
 We again define the degree of controlled synchronization as $\beta = \ln 
-\frac{\tau_c}{\tau}$. For advanced synchronization, $\beta < 0$ indicating 
+(\frac{\tau_c}{\tau}$). For advanced synchronization, $\beta < 0$ indicating 
 that synchronization times upon parameter perturbation is smaller than in 
@@
 We are yet to see if the other technique of generating coherent structures also influence synchronization in these unidirectionally coupled systems.
-\begin{figure}[b]
-	%\includegraphics[scale=0.45]{sync_time_dist_beta.pdf}
-    \includegraphics[scale=0.45]{Fast_strength_dist.eps}
-	\caption{\label{fig:beta_dist}\footnotesize Distribution of degree of 
-	ynchronization, $P(\beta)$, mean is -7.46 and standard deviation 1.67.}
-\end{figure}
 
@@
 which rapidly drives the chaotic trajectories in the sticky neighborhood of 
-regular islands. The distribution of degree of synchronization times indicates 
-the efficiency of the procedure. In both the numerical procedures, we have 
-ignored the complex hierarchical structures which lead to sticky 
-behavior of the chaotic orbits in the phase space. A possible future direction 
-is to investigate the role of stickiness in measure synchronization in coupled 
-Hamiltonian systems.
+regular islands. The distribution of degree of controlled synchronization 
+times indicates the efficiency of the procedure. In both the numerical 
+procedures, we have ignored the complex hierarchical structures which lead to 
+sticky behavior of the chaotic orbits in the phase space. A possible future 
+direction is to investigate the role of stickiness in measure synchronization 
+in coupled Hamiltonian systems.
 

### 2018-09-26 Syncarea_control/paper.tex
paper updated

@@
 
-\title{Mechanism to control synchronization in coupled area-preserving maps}
+\title{Controlling synchronization in coupled area-preserving maps by 
+stickiness}
 
@@
 \begin{abstract}
- A unidirectionally coupled area-preserving maps with mixed phase space may 
+Unidirectionally coupled area-preserving maps with a mixed phase space may 
 show identical synchronization in the sticky neighborhood of the regular 
-islands. We use this fact to devise a couple of numerical procedures to 
-delay and expedite the process of synchronization in standard maps coupled 
-under  Pecora-Caroll coupling scheme. The delay method is based on controlled 
+islands. We use this fact to devise numerical procedures to 
+delay and expedite the process of synchronization in two standard maps coupled 
+under  the Pecora-Caroll coupling scheme. The delay method is based on 
+controlled 
 kicking of trajectories away from synchronization traps for as long as 
-necessary. On the other hand, the method to expiate the process is
+necessary. The method to expiate the process is
 achieved by a parameter perturbation technique which rapidly drives the 
 chaotic trajectories to synchronization traps in the sticky neighborhoods 
-of regular islands. We also point out of the limitations of these methods.
+of regular islands. We also discuss the limitations of these methods.
 \end{abstract}
@@
 \section{Introduction} 
-Low-dimensional Hamiltonian systems commonly exhibit mixed phase space i.e. 
-regular structures and chaotic regions may co-exist at a given degree of 
+A low-dimensional Hamiltonian system commonly exhibits a mixed phase space 
+i.e. regular structures and chaotic regions may co-exist at a given degree of 
 nonlinearity. This mixed nature has interesting consequences for the transport
@@
 of a mixed phase space is the interface between regular regions and chaotic 
-sea \cite{Mackay1984, Easton1993}. The dynamics in these neighbourhoods are 
-complex but fairly well understood \cite{Meiss2015} for lower dimensional 
+sea \cite{Mackay1984, Easton1993}. The dynamics in these neighborhoods are 
+complex but fairly well understood \cite{Meiss2015} for low dimensional 
 systems. The complexity arises from the fact that a chaotic trajectory spends 
@@
 exiting to the chaotic sea. The intermittent tendency of chaotic trajectories 
-to stay close to the regular boundaries is called the phenomenon of 
-stickiness. A major consequence of the stickiness is  the existence of power 
-law in the Poincar{\'e} recurrence times indicating algebraic decay for long 
+to stay close to the regular boundaries is called stickiness. A major 
+consequence of the stickiness is  the existence of power law in the 
+Poincar{\'e} recurrence times indicating algebraic decay for long 
 times times rather than exponential decay expected for normal transport. 
@@
 theoretical level \cite{Altmann2006, Altmann2010,Livorati2012,Bunimovich2012, 
-Kruger2015}. In addition, stickiness has found application in several problems 
-such as particle advection in fluids \cite{Babiano1994,Tel2005}, transport in 
+Kruger2015, Lange2016}. In addition, stickiness has found application in 
+several problems such as particle advection in fluids 
+\cite{Babiano1994,Tel2005}, transport in 
 plasma fusion devices \cite{Szezech2012,Martins2014}, celestial mechanics 
@@
 
-A recent feature of stickiness has emerged from a connected quarter. In our 
-recent work~\cite{Mahata2016}, we have looked at the effects of the mixed 
-phase space on the synchronization in a system of two standard maps coupled in 
+A surprising feature of stickiness has emerged in our earlier 
+work~\cite{Mahata2016}, we have looked at the effects of the mixed phase space 
+on the synchronization in a system of two standard maps coupled in 
 unidirectional drive-response configuration. We have shown that 
@@
 where stickiness have been found to influence synchronization in coupled 
-chaotic systems.  For a chaotic orbit, synchronization typically happens via 
-an intermittent behavior in the phase difference of the drive and response 
-maps.   The sticky neighborhoods of a regular islands contain dynamical traps 
-in chaotic orbits. In such trapping regions in the phase space, parts of a 
-chaotic trajectory are almost regular in time and allow for synchronization to 
-occur. Such traps are characterized using the properties of the finite-time 
-Lyapunov exponent \cite{Szezech2005}. In the context of our work, we will 
-refer to these traps as synchronization traps. We further noted that the 
-behavior of the synchronization traps can be analyzed in a more quantitative 
-way by analyzing the location and stability properties of the periodic orbits 
-in the vicinity of the locations where synchronization takes place.
+chaotic systems. It is important to point out that the role of stickiness in 
+Hamiltonian systems for synchronization was already predicted by 
+Zaslavsky~\cite{Zaslavsky2002b} early in the last decade.  For a chaotic 
+orbit, synchronization typically happens via an intermittent behavior in the 
+phase difference of the drive and response maps.   The sticky neighborhoods of 
+a regular islands {\it temporarily} traps chaotic orbits. In such trapping 
+regions in the phase space, parts of a  chaotic trajectory are almost regular 
+in time and allow for synchronization to occur. Such traps are characterized 
+using the properties of the finite-time Lyapunov exponent \cite{Szezech2005}. 
+In the context of our work, we will refer to these traps as synchronization 
+traps. We further note that the behavior of the synchronization traps can be 
+analyzed in a more quantitative way by analyzing the location and stability 
+properties of the periodic orbits in the vicinity of the locations where 
+synchronization takes place.
 
 Synchronization in coupled chaotic systems, coupled map lattices and networks 
-have been studied in great details over the years.
-Coupled Hamiltonian systems are also known to show synchronization such as measure synchronization~\cite{Hampton1999,Wang2003,Vincent2005,Gupta2017} and identical synchronization~\cite{Mahata2016,Das2017} The fact however remains
+have been studied in details (for reviews, see 
+\cite{Pecora1998,Boccaletti2002, Pecora2015}). 
+Coupled Hamiltonian systems are also known to show synchronization such as 
+measure synchronization~\cite{Hampton1999,Wang2003,Vincent2005,Gupta2017} and 
+identical synchronization~\cite{Mahata2016,Das2017} The fact however remains 
 that the ability to control synchronization based on the dynamics of the 
@@
 area-preserving system of the standard map. We demonstrate the mechanism at $K 
-= 6$ where only two small regular islands exist. It is important to point out 
-that the role of stickiness in Hamiltonian systems for synchronization was 
-already predicted by Zaslavsky~\cite{Zaslavsky2002b} early in the last decade.
+= 6$ where only two small regular islands exist.
 
-The rest of the paper is 
-organized in the following way: we explain the coupling scheme in 
-sec.~\ref{sec:Pecora_Carroll} and the mechanism based on the location of 
-synchronization traps is given in sec.~\ref{Master stability function}. The 
-numerical procedures to delay and advanced synchronization times are 
-demonstrated in sec.~\ref{sec:delay} and sec.~\ref{sec:advanced} respectively, 
-and the paper ends with conclusions in sec.~\ref{sec:conclusions} including 
-discussion on limitations and implications.
+The rest of the paper is organized in the following way: we explain the 
+coupling scheme in sec.~\ref{sec:Pecora_Carroll} and the mechanism based on 
+the location of synchronization traps is given in sec.~\ref{Master stability 
+function}. The numerical procedures to delay and advance synchronization 
+times are demonstrated in Sec.~\ref{sec:delay} and Sec.~\ref{sec:advanced} 
+respectively, and the paper ends with conclusions in 
+Sec.~\ref{sec:conclusions} including discussion on limitations and 
+implications.
+
+\begin{figure}[b]
+    \includegraphics[scale=.45]{Standard_Map}
+    \caption{\label{fig:Standard_map} \footnotesize The phase space of the 
+    standard map for  random initial conditions from the uniform distribution 
+    on $[0,1]$,  at the parameter value K = 6.0.}
+\end{figure}
 
@@
 
-\begin{empheq}[right=\empheqrbrace \mod 1]{align}
+\begin{empheq}[right=\empheqrbrace \mod 1 .]{align}
 P_{n+1} = P_n +  \frac{K}{2\pi}\sin(2\pi Q_n) \nonumber\\
@@
 
-\begin{figure}[b]
-	\includegraphics[scale=.45]{Standard_Map}
-	\caption{\label{fig:Standard_map} \footnotesize The phase space of the standard map for  random initial conditions from the uniform distribution on $[0,1]$,  at the parameter value K = 6.0.}
-\end{figure}
-Here the subscript $n$ denotes the discrete time and $K$ is the nonlinearity parameter.
-These equations typically describe the evolution of two canonical variables $P$ and $Q$ which correspond to the momentum and co-ordinate in the Poincar\'{e} section of a freely moving rotator with interleaved periodic kicks.
-This system represents the behavior of a variety of systems such as charged particle confinement in  mirror magnetic traps, particle dynamics in accelerator, comet dynamics in solar systems etc.
-Two-dimensional  phase space plots  of the standard map for the parameter value $K = 6$ using $25$ initial conditions  are shown in Figs.~\ref{fig:Standard_map}.
+\noindent Here the subscript $n$ denotes the discrete time and $K$ is the 
+nonlinearity parameter. These equations typically describe the evolution of 
+two canonical variables $P$ and $Q$ which correspond to the momentum and 
+co-ordinate in the Poincar\'{e} section of a freely moving rotator with 
+interleaved periodic kicks. This system represents the behavior of a variety 
+of systems such as charged particle confinement in  mirror magnetic traps, 
+particle dynamics in accelerator, comet dynamics in solar systems 
+etc.~\cite{Chirikov1960,Izraelev1980,Chirikov1989,Chirikov2008}
+Two-dimensional  phase space plots  of the standard map for the parameter 
+value $K = 6$ using $25$ initial conditions  are shown in 
+Figs.~\ref{fig:Standard_map}.
 
@@
 	\label{equ:drive}
-	P^d_{n+1} = P^d_n + \frac{K}{2\pi}\sin(2\pi Q^d_n)\nonumber\\
-	Q^d_{n+1} = P^d_{n+1} + Q^d_n \nonumber
+	P^\textnormal{d}_{n+1} = P^\textnormal{d}_n + \frac{K}{2\pi}\sin(2\pi 
+	Q^\textnormal{d}_n)\nonumber\\
+	Q^\textnormal{d}_{n+1} = P^\textnormal{d}_{n+1} + Q^\textnormal{d}_n 
+	\nonumber
 	\end{empheq}
@@
 	\begin{empheq}[right=\empheqrbrace \mod 1]{align}
-	P^r_{n+1} = P^d_n + \frac{K}{2\pi}\sin(2\pi Q^r_n) \nonumber \\
-	Q^r_{n+1} = P^r_{n+1} + Q^r_n \nonumber
+	P^\textnormal{r}_{n+1} = P^\textnormal{d}_n + \frac{K}{2\pi}\sin(2\pi 
+	Q^\textnormal{r}_n) \nonumber \\
+	Q^\textnormal{r}_{n+1} = P^\textnormal{r}_{n+1} + Q^\textnormal{r}_n 
+	\nonumber
 	\end{empheq}
@@
 
-The initial values of $Q$ in the drive and response maps are chosen arbitrarily, whereas the $P$ values are identical.
-The system is said to reach complete synchronization when both the $Q$ values of the drive and response systems evolve asymptotically to identical values i.e.
+The initial values of $Q$ in the drive and response maps are chosen 
+arbitrarily, whereas the $P$ values are identical. The system is said to reach 
+complete synchronization when both the $Q$ values of the drive and response 
+systems evolve identically i.e. 
 \begin{equation}
-\lim_{t\rightarrow\infty}(Q^d_n - Q^r_n) = 0
+\lim_{t\rightarrow\infty}(Q^\textnormal{d}_n - Q^\textnormal{r}_n) = 0
 \end{equation}
@@
 
-In all of the computations reported in the work, two chaotic trajectories are considered to be synchronized to numerical accuracy, if the Euclidean distance between them is less than $10^{-5}$.
-An example is shown in Fig.~\ref{fig:sync_ex}.
-Here, we plot $\Delta Q = Q^d_n-Q^r_n$ for initial conditions $(P^d_0,Q^d_0,Q^r_0) = (0.569, 0.906,0.106)$ at $K = 6$.
-The synchronization time is found to be 570246.
+In all of the computations reported in the work, two chaotic trajectories are 
+considered to be synchronized to numerical accuracy, if the Euclidean distance 
+between them is less than $10^{-5}$. An example is shown in 
+Fig.~\ref{fig:sync_ex}. Here, we plot $\Delta Q = 
+Q^\textnormal{d}_n-Q^\textnormal{r}_n$ for initial 
+conditions $(P^\textnormal{d}_0,Q^\textnormal{d}_0,Q^\textnormal{r}_0) = 
+(0.569, 0.906,0.106)$ at $K = 6$. The synchronization time is found to be 5,70,
+246.
 
@@
 \begin{align}
-\frac{d\overbar{X}_d}{dt} = \overbar{F}(\overbar{X}_d) \nonumber  &&
-\frac{d\overbar{X_r}}{dt} = \overbar{F}(\overbar{X}_r) + \alpha E(\overbar{X}_d-\overbar{X}_r)
+\frac{d\overbar{X}_d}{dt} & = \overbar{F}(\overbar{X}_d) \nonumber  \\
+\frac{d\overbar{X_r}}{dt} &= \overbar{F}(\overbar{X}_r) + \alpha 
+E(\overbar{X}_d-\overbar{X}_r).
 \end{align}
 
-Here $\overbar{X}_d$ and $\overbar{X}_r$ are drive and response variables; the matrix $E$ determines the linear combination of the $\overbar{X}$ used in the difference and $\alpha$ is the coupling strength.
+\noindent Here $\overbar{X}_d$ and $\overbar{X}_r$ are drive and response 
+variables; the matrix $E$ determines the linear combination of the 
+$\overbar{X}$ used in the difference and $\alpha$ is the coupling strength. 
 For the map case, we have the following form
@@
 \begin{align}
-\overbar{X}^d_{n+1} = \overbar{F}(\overbar{X}^d_{n}) \nonumber  &&
-\overbar{X}^r_{n+1} = \overbar{F}(\overbar{X}^r_{n}) + \alpha E(\overbar{X}^d_{n}-\overbar{X}^r_n)
+\overbar{X}^d_{n+1} &= \overbar{F}(\overbar{X}^d_{n}) \nonumber  \\
+\overbar{X}^r_{n+1} &= \overbar{F}(\overbar{X}^r_{n}) + \alpha 
+E(\overbar{X}^d_{n}-\overbar{X}^r_n).
 \end{align}
@@
 P^r_{n+1} &= P^r_n + \frac{K}{2\pi}\sin (2\pi Q^r_n) + \alpha(P^d_n-P^r_n) \nonumber \\
-Q^r_{n+1} &= P^r_{n+1} + Q^r_n 
+Q^r_{n+1} &= P^r_{n+1} + Q^r_n.
 \label{equ:coupled}
@@
 
-To find the stability of the synchronous state, we first express  Eq.(\ref{equ:coupled}) in terms of  $P^\perp = P^d - P^r$ and $Q^\perp = Q^d - Q^r$, as follows
+To find the stability of the synchronous state, we first express  
+Eq.(\ref{equ:coupled}) in terms of  $P^\perp = P^\textnormal{d} - 
+P^\textnormal{r}$ and $Q^\perp = Q^\textnormal{d} -Q^\textnormal{r}$, as 
+follows
 \begin{eqnarray}
 \label{equ:trans}
-P^\perp_{n+1} = (1-\alpha)P^\perp_n+ \frac{K}{2\pi} \sin(2\pi Q^d_n) - \frac{K}{2\pi} \sin(2\pi Q^r_n) \\ \nonumber
-Q^\perp_{n+1} = (1-\alpha)P^\perp_n+ Q^\perp_n + \frac{K}{2\pi} \sin(2\pi Q^d_n) - \frac{K}{2\pi} \sin(2\pi Q^r_n)  
+P^\perp_{n+1} &= (1-\alpha)P^\perp_n+ \frac{K}{2\pi} \sin(2\pi Q^d_n) - 
+\frac{K}{2\pi} \sin(2\pi Q^r_n) \nonumber \\ 
+Q^\perp_{n+1} &= (1-\alpha)P^\perp_n+ Q^\perp_n + \frac{K}{2\pi} \sin(2\pi 
+Q^d_n) \nonumber \\ &- \frac{K}{2\pi} \sin(2\pi Q^r_n).
 \end{eqnarray}
 
-We now write the variational equation for Eq.(\ref{equ:trans}) by linearizing about $(P^d_n,Q^d_n)$
+We now write the variational equation for Eq.(\ref{equ:trans}) by linearizing 
+about $(P^\textnormal{d}_n,Q^\textnormal{d}_n)$
 \begin{equation}
 \label{equ:variational}
-\left[ \begin{array}{c} \delta P^\perp_{n+1} \\ \delta Q^\perp_{n+1} \end{array} \right] = \mathcal{M}(\alpha)\left[ \begin{array}{c}\delta P^\perp_{n} \\ \delta Q^\perp_{n} \end{array} \right]
+\left[ \begin{array}{c} \delta P^\perp_{n+1} \\ \delta Q^\perp_{n+1} 
+\end{array} \right] = \mathcal{M}(\alpha)\left[ \begin{array}{c}\delta 
+P^\perp_{n} \\ \delta Q^\perp_{n} \end{array} \right],
 \end{equation}
@@
 \label{equ:M}
-\mathcal{M}(\alpha) = \begin{bmatrix} 1-\alpha & K\cos(2\pi Q^d_n) \\ 1 - \alpha & 1 + K\cos(2\pi Q^d_n) \end{bmatrix} 
+\mathcal{M}(\alpha) = \begin{bmatrix} 1-\alpha & K\cos(2\pi Q^d_n) \\ 1 - 
+\alpha & 1 + K\cos(2\pi Q^d_n) \end{bmatrix}.
 \end{equation}
 
-This is the master stability equation for the unidirectionally coupled standard map.
-The variational equation (\ref{equ:variational}) is the master stability equation for the coupled system under investigation.
-The associated largest LE computed from the master stability equation is the master stability function of the system, given by:
+This is the master stability equation for the unidirectionally coupled 
+standard map. The variational equation (\ref{equ:variational}) is the master 
+stability equation for the coupled system under investigation. The associated 
+largest Lyapunov exponent computed from the master stability equation is the 
+master stability function of the system, given by:
 \begin{equation}
 \label{equ:msf}
-\lambda  = \lim_{n \rightarrow \infty} \lim_{\delta\overbar{X}_0\rightarrow 0}\frac{1}{n}\sum^{n-1}_{i=0}\ln|JM^n(\overbar{X}_i)|
+\lambda  = \lim_{n \rightarrow \infty} \lim_{\delta\overbar{X}_0\rightarrow 
+0}\frac{1}{n}\sum^{n-1}_{i=0}\ln|JM^n(\overbar{X}_i)|.
 \end{equation}
@@
 \begin{equation} 
-\mathcal{M}(1)= \begin{bmatrix} 0 & K\cos(2\pi Q^d_n) \\ 0 & 1 + K\cos(2\pi Q^d_n) \end{bmatrix} 
+\mathcal{M}(1)= \begin{bmatrix} 0 & K\cos(2\pi Q^d_n) \\ 0 & 1 + K\cos(2\pi 
+Q^d_n) \end{bmatrix},
 \end{equation}
@@
 \label{equ:LE}
-\lambda   = \lim_{n \rightarrow \infty} \frac{1}{n}\sum^{n-1}_{i=0}\ln|1+K\cos(2\pi Q^d_n)|
+\lambda   = \lim_{n \rightarrow \infty} 
+\frac{1}{n}\sum^{n-1}_{i=0}\ln|1+K\cos(2\pi Q^d_n)|.
 \end{equation}
@@
 \label{equ:FLEf}
-\lambda _k(\overbar{X}_0;n)  = \frac{1}{n}\sum^{n-1}_{i=0}\ln|JM^n(\overbar{X}_i)|
+\lambda _k(\overbar{X}_0;n)  = 
+\frac{1}{n}\sum^{n-1}_{i=0}\ln|JM^n(\overbar{X}_i)|.
 \end{equation}
@@
 \begin{equation}
-\lambda_1^s(\overbar{X}_0;n) = \frac{1}{n}\sum^{n-1}_{i=0}\ln|1 + K \cos (2\pi Q_i)|
+\lambda_1^s(\overbar{X}_0;n) = \frac{1}{n}\sum^{n-1}_{i=0}\ln|1 + K \cos (2\pi 
+Q_i)|.
 \end{equation}
 
-The subscript ($k = 1$) has been dropped hereafter as we have only one LE to compute.
+The subscript ($k = 1$) has been dropped hereafter as we have only one 
+exponent to compute.
 
-We plot the variation in the time-5 FTLE defined above at $K = 6$.
-At the point of synchronization, the FTLE values attain a set of negative values consistently in a small window, indicating the existence of a synchronization trap.
-In Fig.~\ref{fig:FTLE} , the three plots show the fluctuations  of FTLE near the point of synchronization for three different sets of initial conditions - $(P^d_0,Q^d_0,Q^r_0) = (0.466,0.098,0.141)$ in (a),  $(0.009,0.026,0.129)$ in (b), and $(0.553,0.554,0.396)$ in (c) with synchronization times 287090, 234521, and 164782 respectively.
-The values on the x-axis indicate time steps at which averages are computed so that synchronization times and the temporal neighborhood are effectively captured in the plots wherein a temporary drop is clearly visible, indicated by ellipses in red.
-Synchronization of chaotic trajectories typically occurs in one of such traps.
+We plot the variation in the time-5 FTLE defined above at $K = 6$. At the 
+point of synchronization, the FTLE values attain a set of smaller values 
+consistently in a small window, indicating the existence of a synchronization 
+trap.  In Fig.~\ref{fig:FTLE} , the three plots show the fluctuations  of FTLE 
+near the point of synchronization for three different sets of initial 
+conditions - $(P^d_0,Q^d_0,Q^r_0) = (0.466,0.098,0.141)$ in (a),  
+$(0.009,0.026,0.129)$ in (b), and $(0.553,0.554,0.396)$ in (c) with 
+synchronization times 287090, 234521, and 164782 respectively. The values on 
+the x-axis indicate time steps at which averages are computed so that 
+synchronization times and the temporal neighborhood are effectively captured 
+in the plots wherein a temporary drop is clearly visible, indicated by 
+ellipses in red. Synchronization of chaotic trajectories typically occurs in 
+one of such traps.
 
@@
 For the numerical procedure, we divide the phase space in $100 \times 100$ 
-grid. The edges of both the islands are detected by applying the standard map 
+grid. The edges of both islands are detected by applying the standard map 
 to point initiated in the chaotic sea, for $10^7$ times.  The edges thus 
 detected in shown in Fig~\ref{fig:edge_reg_island}.  We have chosen the 
-thickness parameter to be $d=0.02$ i.e. euclidean distance of $d$ from points 
+proximity parameter $d=0.02$ i.e. Euclidean distance from points 
 on the edge indicate the extent of the vicinity of the island.  This vicinity, 
@@
 demonstrate this explicitly, we compare the phase angles, defined by $\theta = 
-\tan^{-1}(\frac{Q}{P}$), of the points on the numerically detected edges and 
+\tan^{-1}(\frac{Q}{P}$), of the points on numerically detected edges and 
 the points of synchronization, as 
@@
 \begin{figure}[h]
-	\includegraphics[scale=0.45]{S_U_Percent}
+	\includegraphics[scale=0.42]{S_U_Percent}
 	\caption{\label{fig:Control_success}\footnotesize The curves indicate the 
@@
 
-For the numerical implementation of the mechanism, we have considered $1000$ randomly chosen initial conditions from the chaotic sea of the phase space at $K = 6$.
-The maps have been iterated for $10^7$ times in each case.
-The sticky domain is the determined by the proximity parameter $d_0$ around the numerically detected edges of the regular islands.
-We demonstrate the effectiveness of the control procedure  with respect to several $n-$step controls.
-The computations have been performed for $n \in \mathbb{A}$, where $\mathbb{A} = \{10, 20,30,...,90, 100, 150, ...
-,10 000\}$.
+For the numerical implementation of the mechanism, we have considered $1000$ 
+randomly chosen initial conditions from the chaotic sea of the phase space at 
+$K = 6$. The maps have been iterated for $10^7$ times in each case.  The 
+sticky domain is the determined by the proximity parameter $d_0$ around the 
+numerically detected edges of the regular islands. We demonstrate the 
+effectiveness of the control procedure  with respect to several $n-$step 
+controls. The computations have been performed for $n \in \mathbb{A}$, where 
+$\mathbb{A} = \{10, 20,30,...,90, 100, 150,200, ... ,10 000\}$.
 
 \begin{figure}[t]
-	\includegraphics[scale=0.45]{Sync_time_dist.eps}
+	\includegraphics[scale=0.42]{Sync_time_dist.eps}
 	\caption{\label{fig:Sync_time_dist}\footnotesize Distribution of 
@@
 \begin{figure}[b]
-    \includegraphics[scale=0.45]{Strength_con.eps}
+    \includegraphics[scale=0.42]{Strength_con.eps}
     \caption{\label{fig:Strength_con}\footnotesize Variation of $\beta$ 
@@
 \label{sec:advanced}
-Based on the now explore the possibility to decrease synchronization times 
+We now explore the possibility to decrease synchronization times 
 i.e. to expedite the process of synchronization. Once again, we will make use 
@@
 perturbation technique to generate coherent structures in the phase space of 
-an area-preserving map. However, the technique may also be used to push the 
+an area-preserving map~\cite{Gupte2007}. However, the technique may also be 
+used to push the 
 chaotic bulk on the coherent structure (regular region) already present in the 
@@
 \begin{figure*}[t]
-    \includegraphics[scale=0.5]{Fast_sync_location.eps}
+    \includegraphics[scale=0.45]{Fast_sync_location.eps}
     \caption{\label{fig:location}\footnotesize The phase space of the drive 
@@
         perturbation is applied. Synchronization occurs at the location 
-        indicate 
+        indicated
         by the red `plus' signs.}
@@
 such that $-0.5 \leq P_n \leq 0.5, 0\leq Q_n \leq 1$ . In this form, the 
-standard map is known to have a hyperbolic fixed point at 
-$(0,0)$,$(0,0.5)$. We perturb the parameter $K$ to $K-\epsilon$ if $|P - P_f| 
-< \delta$, $|Q-Q_f| < \delta$. For $P$ and $Q$ outside this $\delta$-strip, 
+standard map has a hyperbolic fixed point at 
+$(P_\textnormal{f},Q_\textnormal{f}) = (0,0)$,$(0,0.5)$. We perturb the 
+parameter $K$ to $K-\epsilon$ if $|P - P_\textnormal{f}| 
+< \delta$, $|Q-Q_\textnormal{f}| < \delta$. For $P$ and $Q$ outside this 
+$\delta$-strip, 
 $K$ does not change.  It may be noted that the perturbed standard map remains 
@@
 for for $K = 6$ , $\delta = 0.4$ and $\epsilon = 2$ for the fixed point 
-$(P_f,Q_f)=(0.0,0.0)$ is shown in Fig.~\ref{fig:location}.   We have also 
+$(P_\textnormal{f},Q_\textnormal{f})=(0.0,0.0)$ is shown in 
+Fig.~\ref{fig:location}.   We have also 
 performed computation for the other fixed point and with other modulo 
 operations as well and results do not differ significantly. Our choice here of 
-modulo operation and the fixed point $(P_f,Q_f)$ has some advantage over the 
+modulo operation and the fixed point $(P_\textnormal{f},Q_\textnormal{f})$ has 
+some advantage over the 
 others in terms of actual numerical values of synchronization times and the 
@@
 response map in accordance with our coupling  scheme described above. We 
-choose a patch around $(P_f,Q_f)$  of length $\delta= 0.4$ which corresponds 
+choose a patch around $(P_\textnormal{f},Q_\textnormal{f})$  of length 
+$\delta= 0.4$ which corresponds 
 to about $16\%$ of the phase space area. We successfully obtain advance 
@@
 \begin{figure}[h]
-    \includegraphics[scale=0.8]{sync_time_beta.eps}
+    \includegraphics[scale=0.7]{sync_time_beta.eps}
     \caption{\label{fig:beta_dist}\footnotesize Distribution of degree of 
-        controlled synchronization, $P(\beta)$, mean is -7.46 and standard 
-        deviation is 1.67.}
+        controlled synchronization, $P(\beta)$, mean is $-7.46$ and standard 
+        deviation is $1.67$.}
 \end{figure}
@@
 \begin{equation}
-f(\beta) =  a\exp\Big(-\frac{(\beta-\mu)^2}{2\sigma^2}\Big)
+f(\beta) =  
+\frac{1}{\sqrt{2\pi\sigma^2}}\exp\Big(-\frac{(\beta-\mu)^2}{2\sigma^2}\Big).
 \end{equation}
@@
 To summarize, we have developed numerical procedures to control 
-synchronization in two identical standard map coupled under the drive-response 
-configuration. The procedure is based on the fact that synchronization in the 
+synchronization in two identical standard maps coupled under the 
+drive-response configuration. The procedure is based on the fact that 
+synchronization in the 
 system typically occurs in the sticky neighborhoods of the regular islands in 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%\bibliographystyle{cpg_unsrt_title}
+\bibliographystyle{cpg_unsrt_title}
 %\bibliographystyle{cpg_unsrt_key_title}
@@
 %\bibliographystyle{unsrt}
-\bibliographystyle{cpg_unsrt_title_for_phys_rev}
+%\bibliographystyle{cpg_unsrt_title_for_phys_rev}
 %\bibliographystyle{cpg_unsrt_key_title2} % to have labels + links

### 2018-09-26 Syncarea_control/paper.tex
paper and ref updated

@@
 
-\documentclass[reprint,superscriptaddress,amsmath,amssymb,aps,pre]{revtex4-1}
+\documentclass[reprint,amsmath,amssymb,aps,pre]{revtex4-1}
 
@@
 Poincar{\'e} recurrence times indicating algebraic decay for long 
-times times rather than exponential decay expected for normal transport. 
+times rather than exponential decay expected for normal transport. 
 Therefore, due to stickiness, even a small regular island can influence the 
@@
 measure synchronization~\cite{Hampton1999,Wang2003,Vincent2005,Gupta2017} and 
-identical synchronization~\cite{Mahata2016,Das2017} The fact however remains 
-that the ability to control synchronization based on the dynamics of the 
-process has not been given much attention, until 
+identical synchronization~\cite{Mahata2016,Das2017} However, the ability to control synchronization based on the dynamics of the process has not been given much attention, until 
 recently~\cite{Grabow2011,Wang2016}. In realistic systems, the speed of 
@@
 to point initiated in the chaotic sea, for $10^7$ times.  The edges thus 
-detected in shown in Fig~\ref{fig:edge_reg_island}.  We have chosen the 
+detected is shown in Fig~\ref{fig:edge_reg_island}.  We have chosen the 
 proximity parameter $d=0.02$ i.e. Euclidean distance from points 
@@
 Fig.~\ref{fig:Edge_Sync_angles}(b) shows the distribution of these phase angles for about 50 000 randomly chosen initial conditions for the drive and response maps that lead to synchronization.
-The locations of the peaks in the distribution shown in (a) approximately matches with this in (b).
+The locations of the peaks in the distribution shown in ~\ref{fig:Edge_Sync_angles}(a) approximately matches with this in ~\ref{fig:Edge_Sync_angles}(b).
 It is, therefore, visibly clear that synchronization occurs in the same angular domain of the phase space where the regular islands exist.
@@
 
-For the numerical implementation of the mechanism, we have considered $1000$ 
+For numerical implementation of the mechanism, we have considered $1000$ 
 randomly chosen initial conditions from the chaotic sea of the phase space at 
 $K = 6$. The maps have been iterated for $10^7$ times in each case.  The 
-sticky domain is the determined by the proximity parameter $d_0$ around the 
+sticky domain is determined by the proximity parameter $d_0$ around the 
 numerically detected edges of the regular islands. We demonstrate the 
@@
 \ln(\frac{\tau_d}{\tau_0}$).  Therefore, for successfully delayed 
-synchronization $\beta>0$ indicating $\tau_c > \tau$.The plot in 
+synchronization $\beta>0$ indicating $\tau_c > \tau$. The plot in 
 Fig.~\ref{fig:Strength_con} shows the variation of $\beta$ i.e. $\beta_{min}$, 
-$\beta_{mean}$, and $\beta_{max}$  indicating minimum, mean and maximum 
+$\beta_{mean}$, and $\beta_{max}$  indicating minimum, mean, and maximum 
 $\beta$ for considered $n-$step controls. Clearly, we are able to attain 
@@
 drive-response configuration. The procedure is based on the fact that 
-synchronization in the 
-system typically occurs in the sticky neighborhoods of the regular islands in 
+synchronization in the system typically occurs in the sticky neighborhoods of the regular islands in 
 the phase space. We have shown that a delay in synchronization can be obtained 
@@
 number of iterations used throughout the computation.  Furthermore, the 
-procedure gives variety of delays. The limitations of this procedure are, one, 
+procedure gives variety of delays. The limitations of this procedure are -- (a) 
 the deflection may insert the drive trajectory in the regular islands which 
-would lead to undesirable faster synchronization and two, a predetermined 
+would lead to undesirable faster synchronization, and (b) a predetermined 
 synchronization time in not obtainable. Furthermore, we have demonstrated a 

### 2018-09-26 Syncarea_control/paper.tex
paper and ref updated
paper updated

@@
 value $K = 6$ using $25$ initial conditions  are shown in 
-Figs.~\ref{fig:Standard_map}.
+Fig.~\ref{fig:Standard_map}.
 
@@
 
-We now analyze synchronization of the coupled system using the master stability function~\cite{Pecora1998}. The numerical procedures to control synchronization times will be discussed thereafter.
+We now analyze synchronization of the coupled system using the master stability function~\cite{Pecora1998}. Numerical procedures to control synchronization times will be discussed thereafter.
 
@@
 \begin{align}
-\frac{d\overbar{X}_d}{dt} & = \overbar{F}(\overbar{X}_d) \nonumber  \\
-\frac{d\overbar{X_r}}{dt} &= \overbar{F}(\overbar{X}_r) + \alpha 
-E(\overbar{X}_d-\overbar{X}_r).
+\frac{d\overbar{X}_\textnormal{d}}{dt} & = \overbar{F}(\overbar{X}_\textnormal{d}) \nonumber  \\
+\frac{d\overbar{X_\textnormal{r}}}{dt} &= \overbar{F}(\overbar{X}_\textnormal{r}) + \alpha 
+E(\overbar{X}_\textnormal{d}-\overbar{X}_\textnormal{r}).
 \end{align}
@@
 \begin{align}
-\overbar{X}^d_{n+1} &= \overbar{F}(\overbar{X}^d_{n}) \nonumber  \\
-\overbar{X}^r_{n+1} &= \overbar{F}(\overbar{X}^r_{n}) + \alpha 
-E(\overbar{X}^d_{n}-\overbar{X}^r_n).
+\overbar{X}^\textnormal{d}_{n+1} &= \overbar{F}(\overbar{X}^\textnormal{d}_{n}) \nonumber  \\
+\overbar{X}^\textnormal{r}_{n+1} &= \overbar{F}(\overbar{X}^\textnormal{r}_{n}) + \alpha 
+E(\overbar{X}^\textnormal{d}_{n}-\overbar{X}^\textnormal{r}_n).
 \end{align}
@@
 \begin{align}
-P^d_{n+1} &= P^d_n + \frac{K}{2\pi}\sin(2\pi Q^d_n) \nonumber\\
-Q^d_{n+1} &= P^d_{n+1} + Q^d_n \nonumber\\
-P^r_{n+1} &= P^r_n + \frac{K}{2\pi}\sin (2\pi Q^r_n) + \alpha(P^d_n-P^r_n) \nonumber \\
+P^\textnormal{d}_{n+1} &= P^\textnormal{d}_n + \frac{K}{2\pi}\sin(2\pi Q^\textnormal{d}_n) \nonumber\\
+Q^\textnormal{d}_{n+1} &= P^\textnormal{d}_{n+1} + Q^\textnormal{d}_n \nonumber\\
+P^\textnormal{r}_{n+1} &= P^\textnormal{r}_n + \frac{K}{2\pi}\sin (2\pi Q^\textnormal{r}_n) + \alpha(P^\textnormal{d}_n-P^\textnormal{r}_n) \nonumber \\
 Q^r_{n+1} &= P^r_{n+1} + Q^r_n.
@@
 \label{equ:trans}
-P^\perp_{n+1} &= (1-\alpha)P^\perp_n+ \frac{K}{2\pi} \sin(2\pi Q^d_n) - 
-\frac{K}{2\pi} \sin(2\pi Q^r_n) \nonumber \\ 
+P^\perp_{n+1} &= (1-\alpha)P^\perp_n+ \frac{K}{2\pi} \sin(2\pi Q^\textnormal{d}_n) - 
+\frac{K}{2\pi} \sin(2\pi Q^\textnormal{r}_n) \nonumber \\ 
 Q^\perp_{n+1} &= (1-\alpha)P^\perp_n+ Q^\perp_n + \frac{K}{2\pi} \sin(2\pi 
-Q^d_n) \nonumber \\ &- \frac{K}{2\pi} \sin(2\pi Q^r_n).
+Q^\textnormal{d}_n) \nonumber \\ &- \frac{K}{2\pi} \sin(2\pi Q^\textnormal{r}_n).
 \end{eqnarray}
@@
 \label{equ:M}
-\mathcal{M}(\alpha) = \begin{bmatrix} 1-\alpha & K\cos(2\pi Q^d_n) \\ 1 - 
+\mathcal{M}(\alpha) = \begin{bmatrix} 1-\alpha & K\cos(2\pi Q^\textnormal{d}_n) \\ 1 - 
 \alpha & 1 + K\cos(2\pi Q^d_n) \end{bmatrix}.
@@
 \begin{equation} 
-\mathcal{M}(1)= \begin{bmatrix} 0 & K\cos(2\pi Q^d_n) \\ 0 & 1 + K\cos(2\pi 
+\mathcal{M}(1)= \begin{bmatrix} 0 & K\cos(2\pi Q^\textnormal{d}_n) \\ 0 & 1 + K\cos(2\pi 
 Q^d_n) \end{bmatrix},
@@
 \lambda   = \lim_{n \rightarrow \infty} 
-\frac{1}{n}\sum^{n-1}_{i=0}\ln|1+K\cos(2\pi Q^d_n)|.
+\frac{1}{n}\sum^{n-1}_{i=0}\ln|1+K\cos(2\pi Q^\textnormal{d}_n)|.
 \end{equation}
@@
 	1 & 1 + (K-\epsilon)\cos(2\pi Q_n)   \\
-	1 & (K-\epsilon)\cos(2\pi Q_n) \end{array} \right)\]
+	1 & (K-\epsilon)\cos(2\pi Q_n) \end{array} \right),\]
 \end{minipage}
@@
 efficiency of the procedure depends upon the number of steps employed in 
-control. By applying the procedure to the system on merely $1\%$ of the phase 
+control. By applying the procedure to the system on almost $16\%$ of the phase 
 space area (at $K = 6$), we have successfully increased synchronization times. 

### 2018-09-28 Syncarea_control/paper.tex
manusript updated + schematic added

@@
 
-\title{Controlling synchronization in coupled area-preserving maps by 
-stickiness}
+\title{Mechanism to controlling synchronization in coupled area-preserving maps}
 
@@
     Stra\ss{}e 38, 01187 Dresden, Germany}
+\email{swetdas@pks.mpg.de}
 
@@
 islands. We use this fact to devise numerical procedures to 
-delay and expedite the process of synchronization in two standard maps coupled 
-under  the Pecora-Caroll coupling scheme. The delay method is based on 
-controlled 
-kicking of trajectories away from synchronization traps for as long as 
-necessary. The method to expiate the process is
-achieved by a parameter perturbation technique which rapidly drives the 
-chaotic trajectories to synchronization traps in the sticky neighborhoods 
+control (delay and expedite) the process of synchronization in two standard maps coupled under  the Pecora-Carroll coupling scheme.
+The delay method is based on controlled kicking of trajectories away from synchronization traps for as long as necessary.
+The method to expedite the process is achieved by a parameter perturbation technique which rapidly drives the chaotic trajectories to synchronization traps in the sticky neighborhoods
 of regular islands. We also discuss the limitations of these methods.
@@
 power law contributions to recurrence and other statistics, and the existence 
-of dynamical traps \cite{Zaslavsky2002a,Zaslavsky2002b}. An interesting region 
+of dynamical traps \cite{Zaslavsky2002a,Zaslavsky2002b}. An intriguing region 
 of a mixed phase space is the interface between regular regions and chaotic 
@@
 in the sticky neighborhood of regular islands in the phase space of the 
-area-preserving system of the standard map. We demonstrate the mechanism at $K 
-= 6$ where only two small regular islands exist.
-
+area-preserving system of the standard map.
+We demonstrate the mechanism at a specific value of the nonlinearly parameter (see Sec.~\ref{Master stability function}) where only two small regular islands exist in the phase space.
+\begin{figure}[t]
+	\includegraphics[scale=.45]{Standard_Map}
+	\caption{\label{fig:Standard_map} \footnotesize The phase space of the standard map at the parameter value K = 6.0.
+A couple of regular islands (shown in the color red) and a connected chaotic region (shown in the color blue).
+}
+\end{figure}
 The rest of the paper is organized in the following way: we explain the 
@@
 
-\begin{figure}[b]
-    \includegraphics[scale=.45]{Standard_Map}
-    \caption{\label{fig:Standard_map} \footnotesize The phase space of the 
-    standard map for  random initial conditions from the uniform distribution 
-    on $[0,1]$,  at the parameter value K = 6.0.}
-\end{figure}
-
-\section{The Pecora Caroll's unidirectional coupling scheme}
+\section{The Pecora Carroll's unidirectional coupling scheme}
 \label{sec:Pecora_Carroll} 
@@
 
+\begin{figure}[t]
+	\includegraphics[scale=.3]{pecora_caroll_schematic.eps}
+	\caption{\label{fig:schematic} \footnotesize Schematic displaying the Pecora-Carroll method of coupling two standard map as explained in the text.}
+\end{figure}
 \noindent Here the subscript $n$ denotes the discrete time and $K$ is the 
@@
 Two-dimensional  phase space plots  of the standard map for the parameter 
-value $K = 6$ using $25$ initial conditions  are shown in 
-Fig.~\ref{fig:Standard_map}.
+value $K = 6$ using a chaotic and a few regular initial conditions  are shown in Fig.~\ref{fig:Standard_map}.
 
@@
 In this case, the $P$ value of the response system is set to the $P$ value of the drive system, at each iterate.
-Therefore, the coupled system is described by the following equations
-
-\begin{minipage}[t]{0.5\textwidth}
-	\begin{empheq}[right=\empheqrbrace \mod 1]{align}\nonumber
-	\label{equ:drive}
-	P^\textnormal{d}_{n+1} = P^\textnormal{d}_n + \frac{K}{2\pi}\sin(2\pi 
-	Q^\textnormal{d}_n)\nonumber\\
-	Q^\textnormal{d}_{n+1} = P^\textnormal{d}_{n+1} + Q^\textnormal{d}_n 
-	\nonumber
-	\end{empheq}
-	\centering \textbf{(Drive)}
-
-\end{minipage}
-\begin{minipage}[t]{0.5\textwidth}
-	\begin{empheq}[right=\empheqrbrace \mod 1]{align}
-	P^\textnormal{r}_{n+1} = P^\textnormal{d}_n + \frac{K}{2\pi}\sin(2\pi 
-	Q^\textnormal{r}_n) \nonumber \\
-	Q^\textnormal{r}_{n+1} = P^\textnormal{r}_{n+1} + Q^\textnormal{r}_n 
-	\nonumber
-	\end{empheq}
-	\centering \textbf{(Response)}
-\end{minipage}
-
-\vspace{.5cm}
+A schematic is given to demonstrate the coupling scheme in Fig.~\ref{fig:schematic}.
 
@@
 \begin{equation}
-\lim_{t\rightarrow\infty}(Q^\textnormal{d}_n - Q^\textnormal{r}_n) = 0
+\lim_{t\rightarrow\infty}(Q^\textnormal{d}_n - Q^\textnormal{r}_n) = 0.
 \end{equation}
@@
 
-In all of the computations reported in the work, two chaotic trajectories are 
-considered to be synchronized to numerical accuracy, if the Euclidean distance 
-between them is less than $10^{-5}$. An example is shown in 
-Fig.~\ref{fig:sync_ex}. Here, we plot $\Delta Q = 
-Q^\textnormal{d}_n-Q^\textnormal{r}_n$ for initial 
+In all of the computations reported in the work, two chaotic trajectories are considered to be synchronized to numerical accuracy, if the Euclidean distance between them is less than $10^{-5}$.
+An example is shown in
+Fig.~\ref{fig:sync_ex}. Here, we plot $\Delta Q = Q^\textnormal{d}_n-Q^\textnormal{r}_n$ for initial 
 conditions $(P^\textnormal{d}_0,Q^\textnormal{d}_0,Q^\textnormal{r}_0) = 
-(0.569, 0.906,0.106)$ at $K = 6$. The synchronization time is found to be 5,70,
-246.
+(0.569, 0.906,0.106)$ at $K = 6$. The synchronization time is found to be 5,70,246.
 
@@
 stability equation for the coupled system under investigation. The associated 
-largest Lyapunov exponent computed from the master stability equation is the 
-master stability function of the system, given by:
+largest Lyapunov exponent (LE) computed from the master stability equation is the master stability function (MSF) of the system, given by:
 \begin{equation}
@@
 A negative value of the MSF (the largest non-zero LE) will ensure that $(P^\perp,Q^\perp) $ tend to zero indicating that the difference between $P$ and $Q$ will die out and the system will synchronize.
-Now, for the Pecora-Carroll approach, we set $\alpha = 1$. This substitution simplifies the matrix in Eq.(\ref{equ:M}) which now reads
+Now, for the Pecora-Carroll approach, we set $\alpha = 1$.
+This substitution simplifies the matrix in Eq.(\ref{equ:M}) which now reads
 
@@
 \lambda   = \lim_{n \rightarrow \infty} 
-\frac{1}{n}\sum^{n-1}_{i=0}\ln|1+K\cos(2\pi Q^\textnormal{d}_n)|.
+\frac{1}{n}\sum^{n-1}_{i=0}\ln|1+K\cos(2\pi Q^\textnormal{d}_i)|.
 \end{equation}
 
-In general, we define the $kth$ time-$n$ Lyapunov exponent associated with an initial point $\overbar{X}_0 = (P_0,Q_0)$ for a map $M(P,Q)$ as
+In general, we define the $kth$ time-$n$ LE associated with an initial point $\overbar{X}_0 = (P_0,Q_0)$ for a map $M(P,Q)$ as
 \begin{equation}
@@
 The synchronization times in each cases are 287090, 234521, and 164782 respectively.
-The time scale on the x-axis indicates iterates steps $t$ at which time-5 FTLEs have been computed.
+The time scale on the $x$-axis indicates iteration steps $t$ at which time-5 FTLEs have been computed.
 The drop window in (a), (b), and (c) shown in ellipses in red occur around synchronization times.
@@
 We extend this notion to the master stability function defined in the Sec.~\ref{Master stability function}  i.e.
-the non-zero Lyapunov exponent defined in Eq.(\ref{equ:LE})
+the non-zero LE defined in Eq.(\ref{equ:LE})
 which takes the following finite time version 
@@
 
-We plot the variation in the time-5 FTLE defined above at $K = 6$. At the 
-point of synchronization, the FTLE values attain a set of smaller values 
-consistently in a small window, indicating the existence of a synchronization 
-trap.  In Fig.~\ref{fig:FTLE} , the three plots show the fluctuations  of FTLE 
-near the point of synchronization for three different sets of initial 
-conditions - $(P^d_0,Q^d_0,Q^r_0) = (0.466,0.098,0.141)$ in (a),  
-$(0.009,0.026,0.129)$ in (b), and $(0.553,0.554,0.396)$ in (c) with 
+We plot the variation in the time-5 finite time Lyapunv exponent (FTLE) defined above at $K = 6$.
+At the point of synchronization, the FTLE values attain a set of smaller values consistently in a small window, indicating the existence of a synchronization trap.
+In Fig.~\ref{fig:FTLE} , the three plots show the fluctuations  of FTLE near the point of synchronization for three different sets of initial conditions - $(P^d_0,Q^d_0,Q^r_0) = (0.466,0.098,0.141)$ in (a), $(0.009,0.026,0.129)$ in (b), and $(0.553,0.554,0.396)$ in (c) with
 synchronization times 287090, 234521, and 164782 respectively. The values on 
-the x-axis indicate time steps at which averages are computed so that 
+the $x$-axis indicate time steps at which averages are computed so that 
 synchronization times and the temporal neighborhood are effectively captured 
@@
 \begin{figure}[b]
-    \includegraphics[scale=0.42]{edge_reg_island.eps}
+    \includegraphics[scale=0.48]{edge_reg_island.eps}
     %\includegraphics[scale=.4]{Standard_Map2.eps}\\
@@
 In Fig.~\ref{fig:Edge_Sync_angles}(a), we show the distribution of phase angles of the points on the edges of regular islands determined by the edge-detection algorithm.
-The bimodality of the distribution is due to the existence of two sharp islands in otherwise chaotic bulk in the phase space.
+Bimodality of the distribution is due to the existence of two sharp islands in otherwise chaotic bulk in the phase space.
 We compare this with the phase angles corresponding to the points in the phase space at which synchronization of the chaotic trajectories occurs.
@@
 		n-step 
-		control (the red curve). The control delays synchronization times. }
+		control (the red curve). The control procedure delays synchronization times. }
 \end{figure}
@@
 Fig.~\ref{fig:Control_success} shows the percentage of successful delays 
-\textbf{(S)} (on the y-axis) for a given $n-$step control (on the x-axis) up 
+\textbf{(S)} (on the y-axis) for a given $n-$step control (on the $x$-axis) up 
 to $n = 5500$ after which the successful cases are always $100\%$.  We achieve 
@@
 number of iterations used throughout the computation.  Furthermore, the 
-procedure gives variety of delays. The limitations of this procedure are -- (a) 
-the deflection may insert the drive trajectory in the regular islands which 
-would lead to undesirable faster synchronization, and (b) a predetermined 
-synchronization time in not obtainable. Furthermore, we have demonstrated a 
-simple way to fasten synchronization by the parameter perturbation technique 
-which rapidly drives the chaotic trajectories in the sticky neighborhood of 
-regular islands. The distribution of degree of controlled synchronization 
-times indicates the efficiency of the procedure. In both the numerical 
-procedures, we have ignored the complex hierarchical structures which lead to 
-sticky behavior of the chaotic orbits in the phase space. A possible future 
-direction is to investigate the role of stickiness in measure synchronization 
-in coupled Hamiltonian systems.
+procedure gives variety of delays.
+The limitations of this procedure are -- (a) the deflection may insert the drive trajectory in the regular islands which would lead to undesirable faster synchronization, and (b) a predetermined synchronization time in not obtainable.
+Furthermore, we have demonstrated a simple way to fasten synchronization by the parameter perturbation technique which rapidly drives the chaotic trajectories in the sticky neighborhood of regular islands.
+The distribution of degree of controlled synchronization times indicates the efficiency of the procedure.
+In both the numerical procedures, we have ignored the complex hierarchical structures which lead to sticky behavior of the chaotic orbits in the phase space.
+A possible future direction is to investigate the role of stickiness in ``measure synchronization" in coupled Hamiltonian systems.
 

### 2018-10-02 Syncarea_control/paper.tex
paper updated

@@
 
-\title{Mechanism to controlling synchronization in coupled area-preserving maps}
+\title{Mechanism to control synchronization in coupled area-preserving maps}
 
@@
 where stickiness have been found to influence synchronization in coupled 
-chaotic systems. It is important to point out that the role of stickiness in 
+chaotic systems. It is important to point out that a possible role of stickiness in 
 Hamiltonian systems for synchronization was already predicted by 
-Zaslavsky~\cite{Zaslavsky2002b} early in the last decade.  For a chaotic 
+Zaslavsky~\cite{Zaslavsky2002b} in beginning of the last decade.  For a chaotic 
 orbit, synchronization typically happens via an intermittent behavior in the 
@@
 The standard map is considered to be the prototypical example of a two-dimensional area-preserving map, and is given by:
-
 \begin{empheq}[right=\empheqrbrace \mod 1 .]{align}
@@
 
-Synchronization time is the least value of the iterate say, $\tau$, where $\Delta Q = Q^d-Q^r$ vanishes and remains so for rest of the iterations. 
+Synchronization time is the least value of the iterate say, $\tau$, where $\Delta Q = Q^d-Q^r$ vanishes for all subsequent iterations: 
 \begin{equation}
-(Q^d_n - Q^r_n) = 0; \hfill  n \geq  \bf {\tau }
+(Q^\textnormal{d}_n - Q^\textnormal{r}_n) = 0; \hfill  n \geq  \bf {\tau }
 \end{equation}
@@
 
-\begin{figure}[t]
+\begin{figure}[b]
 	\includegraphics[scale=.4]{Sync_exmaple}
@@
 \label{Master stability function}
-A general drive-response system coupled unidirectionally may be described by the following set of equations
+A general drive-response system coupled unidirectionally may be described by the following set of equations:
 
@@
 \begin{figure}[b]
-    \includegraphics[scale=0.48]{edge_reg_island.eps}
+    \includegraphics[scale=0.45]{edge_reg_island.eps}
     %\includegraphics[scale=.4]{Standard_Map2.eps}\\
@@
 
-\vspace{0.2cm}
-	where, $ \epsilon \neq 0$ if $|P_f - P_n| < \delta$, $|Q_f - Q_n| < 
+\vspace*{.2cm}
+
+\noindent where, $ \epsilon \neq 0$ if $|P_f - P_n| < \delta$, $|Q_f - Q_n| < 
 	\delta$\\ $\epsilon = 0$, otherwise.\\
@@
 \begin{figure}[h]
-    \includegraphics[scale=0.7]{sync_time_beta.eps}
+    \includegraphics[scale=0.8]{sync_time_beta.eps}
     \caption{\label{fig:beta_dist}\footnotesize Distribution of degree of 
@@
 In both the numerical procedures, we have ignored the complex hierarchical structures which lead to sticky behavior of the chaotic orbits in the phase space.
-A possible future direction is to investigate the role of stickiness in ``measure synchronization" in coupled Hamiltonian systems.
+A possible future direction is to investigate the role of stickiness in measure synchronization in coupled Hamiltonian systems.
+
+\vspace*{0.3cm}
 

### 2018-10-05 Syncarea_control/Das_arxiv/paper.tex
paper updated

--- 2018-10-02 Syncarea_control/paper.tex
@@
 
-\title{Mechanism to control synchronization in coupled area-preserving maps}
+\title{Controlling synchronization in coupled area-preserving maps using 
+stickiness}
 
@@
 
-\pacs{PACS here}
+\pacs{05.45.Xt, 05.45.Ac}
 
@@
 synchronization of chaotic trajectories of the drive and response maps 
-typically occurs in the neighborhood of regular inlands as a consequence of 
+typically occurs in the neighborhood of regular islands as a consequence of 
 stickiness in the region.  This is the first instance, as far as we are aware, 
@@
 analyzed in a more quantitative way by analyzing the location and stability 
-properties of the periodic orbits in the vicinity of the locations where 
+properties of the periodic orbits at the locations where 
 synchronization takes place.
@@
 \begin{figure}[t]
-	\includegraphics[scale=.45]{Standard_Map}
+	\includegraphics[scale=.45]{Standard_Map.png}
 	\caption{\label{fig:Standard_map} \footnotesize The phase space of the standard map at the parameter value K = 6.0.
@@
 \begin{figure}[t]
-	\includegraphics[scale=.3]{pecora_caroll_schematic.eps}
+	\includegraphics[scale=.3]{pecora_caroll_schematic}
 	\caption{\label{fig:schematic} \footnotesize Schematic displaying the Pecora-Carroll method of coupling two standard map as explained in the text.}
@@
 
-We plot the variation in the time-5 finite time Lyapunv exponent (FTLE) defined above at $K = 6$.
-At the point of synchronization, the FTLE values attain a set of smaller values consistently in a small window, indicating the existence of a synchronization trap.
-In Fig.~\ref{fig:FTLE} , the three plots show the fluctuations  of FTLE near the point of synchronization for three different sets of initial conditions - $(P^d_0,Q^d_0,Q^r_0) = (0.466,0.098,0.141)$ in (a), $(0.009,0.026,0.129)$ in (b), and $(0.553,0.554,0.396)$ in (c) with
+We plot the variation in the time-5 finite time Lyapunov exponent (FTLE) 
+defined above at $K = 6$. At the point of synchronization, the FTLE values 
+attain a set of smaller values consistently in a small window, indicating the 
+existence of a synchronization trap.  In Fig.~\ref{fig:FTLE} , the three plots 
+show the fluctuations  of FTLE near the point of synchronization for three 
+different sets of initial conditions - $(P^d_0,Q^d_0,Q^r_0) = 
+(0.466,0.098,0.141)$ in (a), $(0.009,0.026,0.129)$ in (b), and 
+$(0.553,0.554,0.396)$ in (c) with 
 synchronization times 287090, 234521, and 164782 respectively. The values on 
@@
 \begin{acknowledgments}
-I would like to thank Neelima Gupte and Arnd B\"acker for their suggestions to improve the manuscript. 
+I would like to thank Neelima Gupte for discussions and Arnd B\"acker for 
+suggestions to improve the manuscript.
 \end{acknowledgments}
